### 2024-03-21 data_processing/Readme.Rmd
holiday break

---
  title: "RNAseq aCRY processing"
author: "Kelterborn"
date: "2024-03-15"
output:
  md_document:
  variant: gfm
toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

# Download data (AWS CLI)

The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
Index in AWS: bmkdatarelease-37/delivery_20240313120107785
Access key ID: AKIA45O7KHZWJCKW62VE
Secret Access Key: 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
Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
Username:delivery_20240313120107785
Password:Pwd2024@x4S416

```{bash, eval=FALSE}

# install aws
curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
unzip awscliv2.zip
sudo ./aws/install

aws --version

aws configure
AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
AWS Secret Access Key [None]: 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
Default region name [None]: eu-north-1
Default output format [None]: table

fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
cd $fastq

# test
aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt

# all files
aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive

```

# 0. Prepare System
BiocManager::install()

## -R_libraries
```{r library}
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library("ggpubr")
library(vsn)

library(PCAtools)

# library(wget)

```

## -R_folders
```{r r_folders}

```

## -Linux folders
```{bash linux_folders, eval=FALSE}
fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;

dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY"

quantdir="$dir/quants"

indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1"

```

# 1. Mapping
```{bash mapping, eval=FALSE}
mamba activate salmon;

fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);

# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
fastqfiles_all=$fastqfiles;

echo "${fastqfiles:0:100}";
echo "${fastqfiles2:0:100}";
echo "${fastqfiles_all:0:100}";

# files: Unknown_BU327-002T0001_1.fq.gz;
for fn in $fastqfiles_all;
do
bfn=$(echo "`basename ${fn}`");
samp=${bfn%"_1.fq.gz"};
R1=$fn;
R2=$(echo "$R1" | sed 's/_1/_2/');
echo "Processing Sample: $samp";
test -f $R1 && echo "--> File: $R1 exists"
test -f $R2 && echo "--> File: $R2 exists"
outquant=${quantdir}/${samp}_quant
echo $outquant
echo $indexdir

salmon quant -i $indexdir -l A \
  -1 $R1 \
  -2 $R2 \
  -p 25 --validateMappings --gcBias -o $outquant

done

```