### 2024-04-09 1_data_processing/1_data_processing2.Rmd

---
  title: "RNAseq aCRY processing"
author: "Kelterborn"
date: "2024-03-15"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

# 0. Prepare System
BiocManager::install()

## -R_libraries
```{r library}
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library("ggpubr")
library(vsn)

library(PCAtools)

# library(wget)

```

## -R_folders
```{r r_folders}
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
       s <- "S:")
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
list.files(gitdir) %>% head()

fastqdir <- paste(dir,"fastq",sep="/")
list.files(fastqdir) %>% tail()
fastqdir2 <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/",sep="/")
list.files(fastqdir2) %>% head()

indexdir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")

fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
head(readLines(fastaPath,n=50))

quantdir <- paste(dir,"quants",sep="/")

```

## -Linux folders
```{bash linux_folders, eval=FALSE}
fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;

dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY";

quantdir="$dir/quants";

indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";

```

# 1. Raw data processing
## 1a. Download data (AWS CLI)

The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
Index in AWS: bmkdatarelease-37/delivery_20240313120107785
Access key ID: AKIA45O7KHZWJCKW62VE
Secret Access Key: 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
Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
Username:delivery_20240313120107785
Password:Pwd2024@x4S416

```{bash, eval=FALSE}

# install aws
curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
unzip awscliv2.zip
sudo ./aws/install

aws --version

aws configure
AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
AWS Secret Access Key [None]: 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
Default region name [None]: eu-north-1
Default output format [None]: table

fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
cd $fastq

# test
aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt

# all files
aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive

```

### check md5
```{bash}
cd $fastqdir
ls
# 1 file
md5sum Unknown_BU327-002T0008_1.fq.gz

# Check downloads

# Path to the text file containing the MD5 sums
md5_file="data_md5.txt"

# Function to check the MD5 sum of a file
check_md5() {
    file_name=$1
    expected_sum=$2

    md5_sum=$(md5sum "$file_name" | awk '{print $1}')
    if [[ $md5_sum == $expected_sum ]]; then
        echo "$file_name: MD5 sum is correct."
    else
        echo "$file_name: MD5 sum is incorrect."
    fi
}

# Read the MD5 sums from the text file and check them
while IFS= read -r line; do
    file_name=$(echo "$line" | awk '{print $2}')
    expected_sum=$(echo "$line" | awk '{print $1}')
    if [[ -f $file_name ]]; then
        check_md5 "$file_name" "$expected_sum"
    else
        echo "$file_name: File not found."
    fi
done < "$md5_file" | tee -a $log

```

## 1b. Fastqc

## 1c. Mapping
```{bash mapping, eval=FALSE}
mamba activate salmon;

fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);

# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
fastqfiles_all=$fastqfiles;

echo "${fastqfiles:0:100}";
echo "${fastqfiles2:0:100}";
echo "${fastqfiles_all:0:100}";

# files: Unknown_BU327-002T0001_1.fq.gz;
for fn in $fastqfiles_all;
do
bfn=$(echo "`basename ${fn}`");
samp=${bfn%"_1.fq.gz"};
R1=$fn;
R2=$(echo "$R1" | sed 's/_1/_2/');
echo "Processing Sample: $samp";
test -f $R1 && echo "--> File: $R1 exists"
test -f $R2 && echo "--> File: $R2 exists"
outquant=${quantdir}/${samp}_quant
echo $outquant
echo $indexdir

salmon quant -i $indexdir -l A \
  -1 $R1 \
  -2 $R2 \
  -p 25 --validateMappings --gcBias -o $outquant

done

```

# 2. Process Data
## 2a. Sample names
```{r}
# Get Sample table
# Filenames

quantdir <- paste(dir,"quants",sep="/")

f <- list.files(path = quantdir)
files <- as.factor(str_remove(f,pattern ="_quant"))
# short form
sample <- str_remove(files,pattern ="Unknown_")
sample
# condition
library(readxl)
Samplepath <- filePath(path=dir, file="sample_table.xlsx")
Samplefile <- read_xlsx(Samplepath)
Samplefile
samplename <- Samplefile$'Sample name [user defined]'
samplename
genotype <- str_split(samplename, pattern="-", simplify = TRUE)[,1] %>% as.factor() %>% relevel(ref="WT")
genotype

light <- str_split(samplename, pattern="-", simplify = TRUE)[,2] %>% str_remove(pattern="\\d")
# \\d = digits
light <- str_split(samplename, pattern="-", simplify = TRUE)[,2] %>% str_remove_all(pattern="[0-9]") %>% as.factor() %>% relevel(ref="D")
# [0-9]$ number at the end
light

repetition <- str_split(samplename, pattern="-", simplify = TRUE)[,2] %>% str_remove_all(pattern="[A-Z]") %>% as.factor()
repetition

RNA_conc <- Samplefile$'RNA conc. [ng/µl]'
RNA_conc

lane <- str_split(files,pattern ="_",simplify = T)[,5]
lane

condition <- paste(genotype, light, sep="_") %>% factor(levels = c("WT_D","pcry_D","WT_BL","pcry_BL","WT_R","pcry_R"))
condition

# Make table
sample.table <- data.frame(sample.id = sample, filenames = files, sample = samplename, genotype = genotype, treatment = light, condition = condition, repetition = repetition, lane = lane,RNA_conc=RNA_conc)
sample.table
```

## 2b. Mapping Rates

```{r}
library(stringr)
library(R.utils)
library(RColorBrewer)

outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
datadir <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044"
quantdir <- paste(datadir,"quants",sep="/")

list.files(path = outdir)
list.files(path = datadir)
list.files(path = quantdir)

samplelist <- {}
mappingrates <- {}
for (i in list.files(path = quantdir)){
print(i)
si <- paste("",str_sub(i,11,-7),sep = "")
si
samplelist <- c(samplelist,si)
f <- readLines(paste(quantdir,i,"logs/salmon_quant.log", sep="/"))
line <- grep("Mapping rate = ",f,value=TRUE)
sl <- str_length(line)
sl
notime <- substring(line,30,sl)
notime
manual <- substring(line,sl-7,sl-1)
val <- as.numeric(str_extract(notime,"[0-9.]+"))
val
valr<-round(val, digits=2)
print(paste("Mapping rate of ",si," is: ",valr," %"))
mappingrates <- c(mappingrates,valr)
}

# Make table

m.table <- data.frame(sample.table,mappingrates)
```

### Plot mapping rates
```{r}

# Colours

# Plot
plot(m.table$mappingrates)
# -> boring

# increase margin for longer names
par(mar=c(4,15,4,4)+.1)
barplot(height=mappingrates, names=samplename, horiz=T, las=1)
mkdirs("graphs")

pdf(file="graphs/Mapping_Rates.pdf", width=10, height=10)
par(mar=c(4,12,4,4)+.1)
xx <- barplot(height=mappingrates, names=samplename, horiz=T, las=1, xlim=c(0,100)) #col=col
text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.8, col = "red")
dev.off()

```

## 2c. Tximeta

```{r}
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

getwd()
outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
datadir <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044"
quantdir <- paste(datadir,"quants",sep="/")
example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")

list.files(quantdir)[1]
file.exists(example.quant)

# first file:
example.table <- read.table(example.quant, header=T)
head(example.table)
basename(example.quant)
dirname(example.quant)
basename(dirname(example.quant))

# generate file list & prepare variables
files={}
for (i in list.dirs(path = dirname(dirname(example.quant)), full.names = TRUE, recursive = FALSE)) {
files <- c(files,paste(i,"/quant.sf",sep=""))
print(basename(i))
# print(head(read.table(files[length(files)], header=T)))
print (files[length(files)])
}
head(files)

dir <- dirname(files)
head(dir)
head(basename(dir))
run <- basename(dir)
dirdir <- dirname(dir)
head(dirdir)

tximeta_files <- file.path(dirdir, run, "quant.sf") 
file.exists(tximeta_files)
file.exists(files)

# names as working sample names
# see 'm.table' at mapping rates
coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
coldata <- data.frame(coldata,m.table)
coldata
colnames(coldata)
head(coldata$genotype)
condition

# load tximeta
# with linked Transcriptome
## Chlamy
dir1 <- "S:/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy"

indexDir <- file.path(dir1,"chlamy_index_v6.1")
fastaPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
head(readLines(fastaPath,n=50))
gtfPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.repeatmasked_assembly_v6.0.gff3.gz")
readLines(gtfPath,n=20)
gtfPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene_exons.gff3.gz")
readLines(gtfPath,n=20)
gtfPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
readLines(gtfPath,n=20)

file.exists(indexDir, fastaPath, gtfPath)

makeLinkedTxome(indexDir=indexDir,
                source="Phytozome",
                organism="Chlamydomonas reinhardtii",
                release="v6.1",
                genome="CreinhardtiiCC_4532_707",
                fasta=fastaPath,
                gtf=gtfPath,
                write=FALSE)

gtfdata <- readLines(gtfPath)
head(gtfdata, n=20)

# gene_name info sind in gff3 abder nicht in se?

se <- tximeta(coldata, useHub=F)
se

library(SummarizedExperiment)

colData(se)
# meta infos sind da (genotype, treatment,...)

# rename Samples
rownames(colData(se)) <- colData(se)$sample.id

rowRanges(se)
# genome info
seqinfo(se)
# ?
head(assays(se)[["counts"]])
# counts. THE DATA

# Mapping infos:
names(metadata(se)[["quantInfo"]])
str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
metadata(se)[["quantInfo"]]$percent_mapped
metadata(se)[["quantInfo"]]$num_processed

par(mfrow=c(1,2))
barplot(metadata(se)[["quantInfo"]]$percent_mapped, main="Mapping Rate")
barplot(metadata(se)[["quantInfo"]]$num_processed/1000000, main="Mio. Reads")

edb <- retrieveDb(se)
class(edb)
genes(edb)
columns(edb)

se.exons <- addExons(se)
rowRanges(se.exons)[[1]]
gse <- summarizeToGene(se)
rowRanges(gse)

head(assays(gse)[["counts"]])[1:5,1:5]
head(assays(gse)[["abundance"]])[1:5,1:5]
head(assays(gse)[["length"]])[1:5,1:5]

#### add gene symbol
columns(retrieveDb(se))
edb
metadata(se)$txomeInfo$source

TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="first"))
# select()' returned 1:many mapping between keys and columns = 1 gene has many transcripts...
TXNAME.list <- (mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="list"))
head(TXNAME.list)
# info is already there
CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
head(CDSID)
# ?
CDSNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSNAME", keytype = "GENEID", multiVals="first"))
head(CDSNAME)
# same as gene_id
EXONNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "EXONNAME", keytype = "GENEID", multiVals="first"))
head(EXONNAME)
TXTYPE <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXTYPE", keytype = "GENEID", multiVals="first"))
head(TXTYPE)
# no useful info... no gene Symbol!!

# add CDSID
mcols(gse)$CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))

colnames(mcols(gse))
head(rownames(gse))

getwd()
save(gse,file=paste(outdir,"tximeta.txm", sep="/"))
gse <- 1
load(file=paste(outdir,"tximeta.txm", sep="/"))

```