### 2024-04-16 1_data_processing/1_data_processing.Rmd
- separate dds for cia5 & acry - Results Cia5

---
title: "RNAseq aCRY processing"
author: "Kelterborn"
date: "2024-03-15"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

# 0. Prepare System
BiocManager::install()

## -R_libraries
```{r r_libraries}
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(DESeq2)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library("ggpubr")
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(PCAtools)

# library(wget)

```

## -R_folders
```{r r_folders}
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
       s <- "S:")
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
list.files(gitdir) %>% head()

fastqdir <- paste(dir,"fastq",sep="/")
list.files(fastqdir) %>% tail()
fastqdir2 <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/",sep="/")
list.files(fastqdir2) %>% head()

indexDir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")

fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
head(readLines(fastaPath,n=10))

gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
readLines(gtfPath,n=10)

quantdir <- paste(dir,"quants",sep="/")

```

## -Linux folders
```{bash linux_folders, eval=FALSE}
fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;

dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY";

quantdir="$dir/quants";

indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";

```

# 1. Raw data processing
## 1a. Download data (AWS CLI)

The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
Index in AWS: bmkdatarelease-37/delivery_20240313120107785
Access key ID: AKIA45O7KHZWJCKW62VE
Secret Access Key: 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
Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
Username:delivery_20240313120107785
Password:Pwd2024@x4S416

```{bash download data, eval=FALSE}

# install aws
curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
unzip awscliv2.zip
sudo ./aws/install

aws --version

aws configure
AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
AWS Secret Access Key [None]: 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
Default region name [None]: eu-north-1
Default output format [None]: table

fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
cd $fastq

# test
aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt

# all files
aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive

```

### check md5
```{bash md5, eval=FALSE}
cd $fastqdir
ls
# 1 file
md5sum Unknown_BU327-002T0008_1.fq.gz

# Check downloads

# Path to the text file containing the MD5 sums
md5_file="data_md5.txt"

# Function to check the MD5 sum of a file
check_md5() {
    file_name=$1
    expected_sum=$2

    md5_sum=$(md5sum "$file_name" | awk '{print $1}')
    if [[ $md5_sum == $expected_sum ]]; then
        echo "$file_name: MD5 sum is correct."
    else
        echo "$file_name: MD5 sum is incorrect."
    fi
}

# Read the MD5 sums from the text file and check them
while IFS= read -r line; do
    file_name=$(echo "$line" | awk '{print $2}')
    expected_sum=$(echo "$line" | awk '{print $1}')
    if [[ -f $file_name ]]; then
        check_md5 "$file_name" "$expected_sum"
    else
        echo "$file_name: File not found."
    fi
done < "$md5_file" | tee -a $log

```

## 1b. Fastqc

## 1c. Mapping
```{bash mapping, eval=FALSE}
mamba activate salmon;

fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);

# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
fastqfiles_all=$fastqfiles;

echo "${fastqfiles:0:100}";
echo "${fastqfiles2:0:100}";
echo "${fastqfiles_all:0:100}";

# files: Unknown_BU327-002T0001_1.fq.gz;
for fn in $fastqfiles_all;
do
bfn=$(echo "`basename ${fn}`");
samp=${bfn%"_1.fq.gz"};
R1=$fn;
R2=$(echo "$R1" | sed 's/_1/_2/');
echo "Processing Sample: $samp";
test -f $R1 && echo "--> File: $R1 exists"
test -f $R2 && echo "--> File: $R2 exists"
outquant=${quantdir}/${samp}_quant
echo $outquant
echo $indexdir

salmon quant -i $indexdir -l A \
  -1 $R1 \
  -2 $R2 \
  -p 25 --validateMappings --gcBias -o $outquant

done

```

# 2. Process Data
## 2a. Sample names
```{r sample_table}
# Get Sample table
# Filenames

quantdir <- paste(dir,"quants",sep="/")

f <- list.files(path = quantdir)
files <- as.factor(str_remove(f,pattern ="_quant"))
# short form
sample <- str_remove(files,pattern ="Unknown_")
sample
# condition
library(readxl)
Samplepath <- filePath(path=dir, file="sample_table.xlsx")
Samplefile <- read_xlsx(Samplepath)
samplename <- Samplefile$clientName
Samplefile$treatment
Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor(levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red","WT_3HL","WT_16HL","cia5_16HL"))
Samplefile$treatment <- Samplefile$treatment %>% factor(levels = c("dark","blue","red","light","3HL","16HL"))
# Make table
sample.table <- data.frame(Samplefile[,-3])
sample.table %>% kable()

sample.table <- data.frame(lapply(sample.table, factor))

# Check the updated structure of the data frame
str(sample.table)

```

## 2b. Mapping Rates
```{r mapping_rates}

samplelist <- {}
mappingrates <- {}
for (i in list.files(path = quantdir)){
print(i)
si <- paste("",str_sub(i,11,-7),sep = "")
si
samplelist <- c(samplelist,si)
f <- readLines(paste(quantdir,i,"logs/salmon_quant.log", sep="/"))
line <- grep("Mapping rate = ",f,value=TRUE)
sl <- str_length(line)
sl
notime <- substring(line,30,sl)
notime
manual <- substring(line,sl-7,sl-1)
val <- as.numeric(str_extract(notime,"[0-9.]+"))
val
valr<-round(val, digits=2)
print(paste("Mapping rate of ",si," is: ",valr," %"))
mappingrates <- c(mappingrates,valr)
}
sample.table$mappingrates <- mappingrates
# Make table

m.table <- data.frame(sample.table,mappingrates)
```

### Plot mapping rates
```{r plot_mappingrates}

# Plot
par(mar=c(2,4,0,0)+.1) # set margins for figure
plot(sample_table_all$mappingrates)
# -> boring

# increase margin for longer names
par(mar=c(2,5,2,2)+.1)
# barplot(height=mappingrates, names=samplename, horiz=T, las=1)

xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename, horiz=T, las=1, xlim=c(0,100))
text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.5, col = "red")

par(mar=c(4,4,4,4)+.1)

```

## 2c. Tximeta

```{r tximeta}
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

getwd()
example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")

list.files(quantdir)[1]
file.exists(example.quant)

# first file:
example.table <- read.table(example.quant, header=T)
head(example.table)
basename(example.quant)
dirname(example.quant)
basename(dirname(example.quant))

# generate file list & prepare variables
files={}
for (i in list.dirs(path = dirname(dirname(example.quant)), full.names = TRUE, recursive = FALSE)) {
files <- c(files,paste(i,"/quant.sf",sep=""))
# print(basename(i))
# print(head(read.table(files[length(files)], header=T)))
# print (files[length(files)])
}
head(files)

dirfiles <- dirname(files)
head(dirfiles)
head(basename(dirfiles))
run <- basename(dirfiles)
dirdir <- dirname(dirfiles)
head(dirdir)

tximeta_files <- file.path(dirdir, run, "quant.sf") 
file.exists(tximeta_files)
file.exists(files)

# names as working sample names
# see 'm.table' at mapping rates
coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
coldata <- data.frame(coldata,sample.table)
coldata %>% head() %>% kable()

# load tximeta
# with linked Transcriptome
## Chlamy
file.exists(indexDir, fastaPath, gtfPath)

makeLinkedTxome(indexDir=indexDir,
                source="Phytozome",
                organism="Chlamydomonas reinhardtii",
                release="v6.1",
                genome="CreinhardtiiCC_4532_707",
                fasta=fastaPath,
                gtf=gtfPath,
                write=FALSE)

gtfdata <- readLines(gtfPath)
head(gtfdata, n=10)

# gene_name info sind in gff3 abder nicht in se?
coldata_Cre <- filter(coldata, species == "Cre")
se <- tximeta(coldata_Cre, useHub=T)
se

library(SummarizedExperiment)

colData(se) %>% head() %>% kable()
# meta infos sind da (genotype, treatment,...)

# rename Samples
rownames(colData(se)) <- colData(se)$clientName

rowRanges(se) %>% kable()
# genome info
seqinfo(se)
# ?
head(assays(se)[["counts"]]) %>% kable()
# counts. THE DATA

# Mapping infos:
names(metadata(se)[["quantInfo"]])
# str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
metadata(se)[["quantInfo"]]$percent_mapped
metadata(se)[["quantInfo"]]$num_processed

par(mfrow=c(1,2))
barplot(metadata(se)[["quantInfo"]]$percent_mapped, main="Mapping Rate")
barplot(metadata(se)[["quantInfo"]]$num_processed/1000000, main="Mio. Reads")
par(mfrow=c(1,1))

edb <- retrieveDb(se)
class(edb)
genes(edb)
columns(edb)

se.exons <- addExons(se)
rowRanges(se.exons)[[1]]
gse <- summarizeToGene(se)
rowRanges(gse)

head(assays(gse)[["counts"]])[1:5,1:5]
head(assays(gse)[["abundance"]])[1:5,1:5]
head(assays(gse)[["length"]])[1:5,1:5]

#### add gene symbol
columns(retrieveDb(se))
edb
metadata(se)$txomeInfo$source

TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="first"))
# select()' returned 1:many mapping between keys and columns = 1 gene has many transcripts...
TXNAME.list <- (mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="list"))
head(TXNAME.list)
# info is already there
CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
head(CDSID)
# ?
CDSNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSNAME", keytype = "GENEID", multiVals="first"))
head(CDSNAME)
# same as gene_id
EXONNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "EXONNAME", keytype = "GENEID", multiVals="first"))
head(EXONNAME)
TXTYPE <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXTYPE", keytype = "GENEID", multiVals="first"))
head(TXTYPE)
# no useful info... no gene Symbol!!

# add CDSID
mcols(gse)$CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))

colnames(mcols(gse))
head(rownames(gse))

getwd()
# save(gse,file=paste(dir,"tximeta.RDS", sep="/"))
# gse <- 1
# load(file=paste(dir,"tximeta.RDS", sep="/"))

```

### -Add gene info from Master annotation

Use other files from Phytozome database: CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv

```{r anno}
library(data.table)
library(plyr)
library(tidyverse)
library(ape)

# more gene info from master_annotation_table
annopath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")

anno <- read.delim2(annopath, header = T, sep = "\t",)
head(anno)[1:5,1:9] %>% kable()

# search for genes
anno[str_detect(anno[["locusName_4532"]],"Cre01.g000050"),]
anno[str_detect(anno[["geneSymbol"]],"HKR"),]
anno[str_detect(anno[["previousIdentifiers"]],"HKR"),]
anno[str_detect(anno[["geneSymbol"]],"COP"),]
anno[str_detect(anno[["previousIdentifiers"]],"COP1"),]
anno[str_detect(anno[["geneSymbol"]],"cop"),]
anno[str_detect(anno[["geneSymbol"]],"PLAP6"),]
anno[str_detect(anno[["geneSymbol"]],"COQ3"),]
anno[str_detect(anno[["geneSymbol"]],"PCRY"),]
anno[str_detect(anno[["geneSymbol"]],"CRY"),]
anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH"),]

dim(anno)

# Remove "_4532"
anno$gene_id <- str_remove(anno$locusName_4532,pattern = "_4532")
rownames(anno) <- anno$gene_id

# convert previousIdentifiers to list
anno$previousIdentifiers_list <- strsplit(anno$previousIdentifiers, "#")
head(anno)
anno$previousIdentifiers_list[1]

# polish "previousIdentifiers"
# remove Cre & g-numbers
head(anno$previousIdentifiers)
tail(anno$previousIdentifiers)
prev.symbols <-  
  str_replace_all(anno$previousIdentifiers,pattern="Cre\\d+.g\\d+.t\\d+.\\d+", replacement="?") %>% 
  str_replace_all(pattern="Cre...g\\d+", replacement="?") %>% 
  str_replace_all(pattern="g\\d+.t\\d+", replacement="?") %>% 
  str_remove_all(pattern="\\?") %>%
  str_remove_all("^#|#$| ") %>% str_remove_all("^#|#$") %>% str_remove_all("^#|#$|^0") %>% 
  str_replace_all(pattern="##","#") %>% str_replace_all(pattern="##","#")
prev.symbols[str_detect(prev.symbols,"#")] %>% head(n=20)
prev.symbols[str_detect(prev.symbols,"0")] %>% head(n=20)
prev.symbols[str_detect(prev.symbols,"Cre")] %>% head(n=20)
tail(anno$previousIdentifiers[str_detect(anno$previousIdentifiers,"Cre")])

anno$prev.symbols <- prev.symbols
head(anno)[1:5,1:9] %>% kable()

# Make colum with gene symbol or prev.symbol or Gene id
anno$combname <- anno$geneSymbol

summary(anno$combname)
summary(anno$combname=="")
summary(anno$combname==" ")
summary(is.na(anno$combname))
summary(anno$combname=="0")

# replace "" with prev.symbol
summary(anno$combname=="")
  length(anno$combname[anno$combname==""])
  length(str_split(
  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1])
anno$combname[anno$combname==""] <- str_split(
  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1]
summary(anno$combname=="")

anno$combname[anno$combname==""] <- anno$gene_id[anno$combname==""]
summary(anno$combname=="")
subset(anno,anno$combname=="")

summary(duplicated(anno$combname))
# some duplicates

colnames(anno)[which(names(anno) == "combname")] <- "id.symbol"

summary(anno$geneSymbol!="")
# 4408 gene Symbols
summary(anno$previousIdentifiers!="")
# all (!) 17712 genes have "previousIdentifiers"
summary(anno$prev.symbols!="")
# 7236 have alternative/old symbols
summary(anno$id.symbol)

dim(anno)

# anno[1:10,1:10]

# save(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"anno.RDS",sep="/"))

```

### -add anno data into gse ####
```{r anno_gse}

# load(file=paste(dir,"tximeta.RDS", sep="/"))
# load(file=paste(dir,"/anno.RDS", sep="/"))
anno[1:3,1:8] %>% kable()

# Remove "_4532"
mcols(gse)$gene_id2 <- mcols(gse)[,"gene_id"]
mcols(gse)$gene_id <- str_remove(mcols(gse)[,"gene_id2"], pattern = "_4532")
rownames(gse) <- mcols(gse)$gene_id

summary(anno$gene_id %in% mcols(gse)$gene_id)
summary(mcols(gse)$gene_id %in% anno$gene_id)

# choose info
gsetable <- as.data.frame(mcols(gse)[,c("gene_id","gene_id2")])

anno_join <- anno[,c("gene_id", "geneSymbol", "id.symbol","prev.symbols", "previousIdentifiers", "previousIdentifiers_list", "Description", "Comments", "TargetP", "Predalgo")]
anno_join %>% head() %>% kable()

class(anno_join)
dim(anno_join)

# any duplicated:
summary(duplicated(anno$gene_id))

new <- plyr::join(gsetable,anno_join, by = "gene_id", type = "left")
dim(new)
head(new) %>% kable()

new2 <- left_join(gsetable,anno_join,by = "gene_id")
dim(new2)
head(new2) %>% kable()

mcols(gse) <- left_join(as.data.frame(mcols(gse)),anno_join,by = "gene_id")
mcols(gse) %>% head() %>% kable()

save(gse,file=paste(dir,"tximeta.RDS",sep="/"))
```

## 2c. DESeq2 Analysis

```{r deseq2}
getwd()
# load(file=paste(dir,"tximeta.RDS", sep="/"))

colData(gse) %>% head() %>% kable()
mcols(gse) %>% head() %>% kable()

gse_acry <- gse[,colData(gse)$experiment=="acry"]
colData(gse_acry) <- colData(gse_acry) %>% droplevels()
colData(gse_acry)$genotype <- colData(gse_acry)$genotype %>% relevel(ref="WT")

gse_cia5 <- gse[,colData(gse)$experiment=="cia5"]
colData(gse_cia5) <- colData(gse_cia5) %>% droplevels()
colData(gse_cia5)$genotype <- colData(gse_cia5)$genotype %>% relevel(ref="WT")

```

### Deseq2 acry
```{r}
design <- ~condition
# design <- ~genotype+treatment+genotype:treatment

dds <- DESeqDataSet(gse_acry, design=design)
colData(dds) %>% head() %>% kable()

##########################################\n
## filter all rows with rowsum = 0 ####
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

# min counts
# at least 5 counts in 3 samples
keep.sn <- rowSums(counts(dds) >= 5) >= 3
keep.sn %>% summary()
dds <- dds[keep.sn,]

dds <- estimateSizeFactors(dds)
dds
head(assays(dds)[["counts"]])[1:5,1:5]
length(rownames(counts(dds)))

colData(dds) %>% head() %>% kable()
# add metadata
# colData(dds)$strain <- factor(colData(dds)$strain)

# ## combine technical replicates #####
# colData(dds)
# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
# colData(dds)
# dds$runsCollapsed
# colnames(dds)
# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))

###########
# run DESeq
dds <- DESeq(dds)
DESeq2::plotMA(dds)

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition")

head(mcols(dds)$geneSymbol)

anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
# no hit
g <- anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH1"),"gene_id"]
g
plotCounts(dds, gene = g, intgroup = "condition")
g <- anno[str_detect(anno[["geneSymbol"]],"PCRY"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
g <- anno[str_detect(anno[["geneSymbol"]],"ACRY"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
g <- anno[str_detect(anno[["geneSymbol"]],"ROC15"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
g <- anno[str_detect(anno[["geneSymbol"]],"ROC40"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])

# search genes
anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()

anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()

###############
# save dds #####

# save(dds, file = paste(dir,"dds_23design.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

```

### Deseq2 cia5
```{r}
design <- ~genotype
dds <- DESeqDataSet(gse_cia5, design=design)
colData(dds) %>% head() %>% kable()

## filter all rows with rowsum = 0 ####
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,4000), xlim = c(0,10))
hist(counts(dds), breaks=300000, ylim = c(0,10000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

# min counts
# at least 5 counts in 3 samples
keep.sn <- rowSums(counts(dds) >= 5) >= 3
keep.sn %>% summary()
dds <- dds[keep.sn,]

dds <- estimateSizeFactors(dds)
dds
head(assays(dds)[["counts"]])[1:5,1:5]
length(rownames(counts(dds)))

colData(dds) %>% head() %>% kable()

# run DESeq
dds <- DESeq(dds)

# qc check
DESeq2::plotMA(dds)
plotDispEsts(dds)

# counts
plotCounts(dds, gene = "Cre01.g000150", intgroup = "genotype", col=colData(dds)$genotype)

save(dds, file = paste(dir,"dds_cia5.RDS",sep="/"), compress = FALSE)

```

# 3. Pre-Analysis
### load data
```{r}
load(file=paste(dir,"dds_cia5.RDS", sep="/"))
load(file=paste(dir,"dds_acry.RDS", sep="/"))
load(file=paste(dir,"anno.RDS", sep="/"))

```

### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
ntd <- normTransform(dds)
rld <- rlog(dds, blind=FALSE) #regularized logarithm
```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="33%"}
# load(file=paste(data,"rlog.rld", sep="/"))
meanSdPlot(assay(ntd))
meanSdPlot(assay(vsd))
meanSdPlot(assay(rld))

```

### - Check sample distance
```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype")])
anno_colors <- list(treatment = c("grey20","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                    genotype = c("grey","seagreen3","tan2"))

names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         annotation_col=anno_col,
         annotation_colors = anno_colors,
         show_colnames     = FALSE,
         col=viridis(20),
         cutree_rows = 6,
         cutree_cols = 6,
         fontsize_row = 5)

```

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))
vst_dat <- assay(vst(dds))

### PCA with top 500 genes with highest row variance 
pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE)
# levels(pcaData$genotype)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "treatment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment)

g2 <- ggplot(pcaData, aes(PC1, PC2, color=genotype, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "genotype") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = anno_colors$genotype)

g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "condition") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  # coord_fixed()
  scale_color_viridis_d(option ="viridis")

((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")

```

###### -- Advanced PCA
```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# calculate PCA (all data)
## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots

# p <- pca(vst_dat, metadata = colData(dds))
# vst <- assay(vst(dds))
p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)

# check different PCAs
p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment

p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype

p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition

(p1+p2) / (p4)

# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

# plot Scree
# find explaining PCs
# horn <- parallelPCA(vst_dat)
# horn$n = 2
# warnings()

horn <- list()
horn$n <- 2

screeplot(p,
    components = getComponents(p, 1:20),
    vline = c(horn$n, elbow)) +
    geom_label(aes(x = horn$n + 1, y = 30,
      label = 'Horn\'s=2', vjust = -1, size = 8)) +
    geom_label(aes(x = elbow + 1, y = 50,
      label = 'Elbow=4', vjust = -1, size = 8))

bi <- biplot(p,x="PC3",y="PC1",
    lab = p$metadata$condition,
    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    labSize = 3, legendIconSize = 4.0,
    legendPosition = 'bottom',
    sizeLoadingsNames = 3,
    axisLabSize = 10,
    captionLabSize = 1)

bi <- bi +theme(panel.background = element_rect(fill='transparent'))

pairs <- pairsplot(p,
    components = getComponents(p, c(1:7)),
    triangle = TRUE, trianglelabSize = 6,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),
    plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

layout <- c(
  area(t = 1, l = 1, b = 6, r = 6),
  area(t = 5, l = 1, b = 7, r = 3)
)
pairs + bi + 
  plot_layout(design = layout)

# pairs + inset_element(bi, left = 0, bottom = 0, right = 0.5, top = 0.5)

plotloadings(p,
             components = getComponents(p)[1:7],
             col = viridis(3),
             labSize = 3)

```

### - Plot example counts
#### cia5

```{r example_counts, fig.show="hold", out.width="50%", eval=FALSE, include=TRUE, echo=FALSE}

load(file=paste(dir,"dds_cia5.RDS", sep="/"))

## cia5

# plap6
PLAP6 <- "Cre03.g188700"
d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)

resultsNames(dds)
res <- results(dds,contrast = c("genotype","cia5","WT"))
res["Cre03.g188700",]
ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
   # scale_fill_manual(values=cols[c(1,5)]) +
  #  scale_color_manual(values=cols[c(1,5)]) +
   scale_y_continuous(trans = "log2") +
  scale_fill_manual(values = c("grey30","salmon")) + 
  scale_color_manual(values = c("grey30","salmon")) + 
    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
  assign(paste("gcounts_",ig,sep=""),gcounts)

  gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
# ggsave(paste(outdir,"graphs/2023_11_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
#   width = 18,
#   height = 10)

goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "DCRY2", "PHOT1", "UVR8", "HKR1") #"HKR2"

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Get Count data
i <- 1
l <- length(goi_phot)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=anno_phot[i,"gene_id"], intgroup=c("condition","genotype","treatment"), main=anno_phot[i,"id.symbol"],returnData=TRUE)
  d$Gene <- rep(anno_phot[i,"id.symbol"],length(rownames(d)))
  d$sample <- rownames(d)
  # rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "PHOT1", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","treatment","genotype"),col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene, levels = anno_tf$geneSymbol[c(9,6,8,7,15,2,1,4,3,5,13,11,14,10,12)])
levels(all_counts$Gene)

# colours
# group.colors <- rep(anno_colors$treatment, each = 2)
group.colors <- rep(anno_colors$treatment, each = 2)[c(1:6,7,9,11)]
names(group.colors) <- colData(dds)$condition %>% levels()
anno_colors$condition <- group.colors

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts

cols = brewer.pal(n=9,name = 'Paired')

gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition, color=condition)) +
  geom_boxplot(color="black") +
  geom_point(position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  scale_y_continuous(trans = "log2")
gcounts

gcounts2 <- ggplot(all_counts, aes(x = condition, y = count, fill=Gene)) +
  geom_boxplot(fatten = 1) +
  scale_fill_viridis_d() +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts3 <- ggplot(all_counts[all_counts$Gene=="ROC40",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
  geom_boxplot(color="black") +
  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=anno_colors$treatment) +
  scale_color_manual(values=anno_colors$treatment) +
  scale_y_continuous(trans = "log2") +
  labs(title = "ROC40") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
gcounts3

gcounts4 <- ggplot(all_counts[all_counts$Gene=="ROC114",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
  geom_boxplot(color="black") +
  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=anno_colors$treatment) +
  scale_color_manual(values=anno_colors$treatment) +
  scale_y_continuous(trans = "log2") +
  labs(title = "ROC114") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
gcounts4

```

```{r}
sessionInfo()
```