### 2024-04-16 2_A_Results_cia5/2_results_cia5.Rmd

---
title: "RNAseq cia5 results"
author: "Kelterborn"
date: "2024-04-16"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

# 0. Prepare System
BiocManager::install()

## -R_libraries
```{r r_libraries}
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(DESeq2)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library("ggpubr")
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(PCAtools)

# library(wget)

```

## -R_folders
```{r r_folders}
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
       s <- "S:")
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
list.files(gitdir) %>% head()

fastqdir <- paste(dir,"fastq",sep="/")
list.files(fastqdir) %>% tail()
fastqdir2 <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/",sep="/")
list.files(fastqdir2) %>% head()

indexDir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")

fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
head(readLines(fastaPath,n=10))

gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
readLines(gtfPath,n=10)

quantdir <- paste(dir,"quants",sep="/")

```

# 1. Load data
```{r load_data}
load(file=paste(dir,"dds_cia5.RDS", sep="/"))
load(paste(dir,"anno.RDS",sep="/"))

```

### - Plot counts
#### -- cia5 & rbcs

```{r example_counts, fig.show="hold", out.width="50%", eval=TRUE, include=TRUE, echo=FALSE}

## cia5

# plap6
PLAP6 <- "Cre03.g188700"
d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)
g1 <- ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
   scale_y_continuous(limits = c(60,150),trans = "log2") +
  scale_fill_manual(values = c("grey30","salmon")) + 
    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))

# rbcl
anno[str_detect(anno[["Description"]],"RuBisCO"),1:9]
rbcl <- "CreCp.g802313"
d <- plotCounts(dds,gene = rbcl, intgroup= c("treatment","genotype","condition"), returnData=TRUE)
g2 <- ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
   scale_y_continuous(trans = "log2") +
  scale_fill_manual(values = c("grey30","salmon")) + 
#  scale_color_manual(values = c("grey30","salmon")) + 
    labs(title = paste("rbcl (",rbcl,")",sep=""))
g1
g2
g1+g2 + plot_layout(guides = "collect", axes="collect", axis_titles="collect") & theme(legend.position = 'right')

```

# 2. Results
```{r results}
resultsNames(dds)
res <- results(dds,contrast = c("genotype","cia5","WT"))
res["Cre03.g188700",]

```

```{r}
sessionInfo()
```