### 2024-04-20 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
acry pcry combine

---
title: "RNAseq aCRY pCRY data dive"
author: "Kelterborn"
date: "2024-04-19"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'Readme.md')
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

BiocManager::install("ggalt") # , type="source"

# Load System
## -R_libraries
```{r r_libraries}
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(DESeq2)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library(ggpubr)
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)

# library(wget)

```

## -R_folders
```{r r_folders}

# Linux Workstation
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
    ifelse(Sys.info()["sysname"]== "Darwin",
       s <- "/mnt/s",
       s <- "S:"))
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
# list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
# list.files(gitdir) %>% head()

# Macbook
scriptdir <- rstudioapi::getSourceEditorContext()$path
gitdir <- scriptdir %>% dirname() %>% dirname()
dir <- gitdir %>% dirname()

```

# 1. Load aCRY & pCRY
## - Tximeta files
```{r get_tximeta_files}

# Annotation file
load(file=paste(dir,"anno.RDS", sep="/"))

# Seq 1
tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm"
load(file=tximeta_pcry_file)
gse_pcry <- gse

# Seq 2 (acry + cia5)
load(file=paste(dir,"tximeta.RDS", sep="/"))
gse_acry <- gse[,colData(gse)$experiment=="acry"]
colData(gse_acry) <- colData(gse_acry) %>% droplevels()
colData(gse_acry)$genotype <- colData(gse_acry)$genotype %>% relevel(ref="WT")

# combine
( mcols(gse_acry) == mcols(gse_pcry) ) %>% summary()
# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
names(colData(gse_pcry)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
colData(gse_acry) <- colData(gse_acry)[,c(1:4,7:12)]
colData(gse_pcry)$experiment <- "pcry"
colData(gse_pcry) <- colData(gse_pcry)[,c(1,3,2,4:6,8,12,7,11)]  
# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()

# now same metadata
gse <- cbind(gse_acry,gse_pcry)

colData(gse)$genotype

colData(gse)$treatment <- colData(gse)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
  str_replace(pattern="BL", replacement = "blue") %>%
  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))

colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","pcry_dark","WT_blue","acry_blue","pcry_blue","WT_red","acry_red","pcry_red"))

colData(gse)$experiment

```

## Run Deseq2
```{r dds_acry_pcry}
# design <- ~condition
design <- ~genotype+treatment+genotype:treatment

dds <- DESeqDataSet(gse, design=design)
colData(dds) %>% head() %>% kable()

##########################################\n
# filter all rows with rowsum = 0 ####\n
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

# min counts
# at least 5 counts in 3 samples
keep.sn <- rowSums(counts(dds) >= 5) >= 3
keep.sn %>% summary()
dds <- dds[keep.sn,]

dds <- estimateSizeFactors(dds)
dds
head(assays(dds)[["counts"]])[1:5,1:5]
length(rownames(counts(dds)))

# colData(dds) %>% head() %>% kable()
# add metadata
# colData(dds)$strain <- factor(colData(dds)$strain)

# ## combine technical replicates #####\n
# colData(dds)
# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
# colData(dds)
# dds$runsCollapsed
# colnames(dds)
# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))

###########\n
# run DESeq
dds <- DESeq(dds)

plotDispEsts(dds)
resultsNames(dds)
DESeq2::plotMA(dds)

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$experiment)

head(mcols(dds)$geneSymbol)

###############\n
# save dds #####

# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

```

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
ntd <- normTransform(dds)
rld <- rlog(dds, blind=FALSE) #regularized logarithm
```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
# load(file=paste(data,"rlog.rld", sep="/"))
meanSdPlot(counts(dds, normalized =TRUE))
meanSdPlot(assay(ntd))
meanSdPlot(assay(vsd))
meanSdPlot(assay(rld))

```

### - Check sample distance
```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                    genotype = c("green4","orchid","tan2"))

names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         annotation_col=anno_col,
         annotation_colors = anno_colors,
         show_colnames     = FALSE,
         col=viridis(20),
         cutree_rows = 6,
         cutree_cols = 6,
         fontsize_row = 5)

```

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))
vst_dat <- assay(vst(dds))

### PCA with top 500 genes with highest row variance 
pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE)
# levels(pcaData$genotype)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "treatment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment)

g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "experiment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = anno_colors$genotype)

g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "condition") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  # coord_fixed()
  scale_color_viridis_d(option ="viridis")

((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")

```

###### -- Advanced PCA
```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# calculate PCA (all data)
## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots

# p <- pca(vst_dat, metadata = colData(dds))
# vst <- assay(vst(dds))
p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)

# check different PCAs
p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment

p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype

p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition

(p1+p2) / (p4)

# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

# plot Scree
# find explaining PCs
horn <- parallelPCA(vst_dat)
# horn$n = 2
# warnings()

# horn <- list()
# horn$n <- 2

screeplot(p,
    components = getComponents(p, 1:20),
    vline = c(horn$n, elbow)) +
    geom_label(aes(x = horn$n + 1, y = 30,
      label = paste('Horn',horn$n,sep="="), vjust = -1, size = 8)) +
    geom_label(aes(x = elbow + 1, y = 50,
      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8))

bi <- biplot(p,x="PC3",y="PC1",
    lab = p$metadata$condition,
    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    labSize = 3, legendIconSize = 4.0,
    legendPosition = 'bottom',
    sizeLoadingsNames = 3,
    axisLabSize = 10,
    captionLabSize = 1)

bi <- bi +theme(panel.background = element_rect(fill='transparent'))

pairs <- pairsplot(p,
    components = getComponents(p, c(1:7)),
    triangle = TRUE, trianglelabSize = 6,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),
    plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

layout <- c(
  area(t = 1, l = 1, b = 6, r = 6),
  area(t = 5, l = 1, b = 7, r = 3)
)
pairs + bi + 
  plot_layout(design = layout)

# pairs + inset_element(bi, left = 0, bottom = 0, right = 0.5, top = 0.5)

plotloadings(p,
             components = getComponents(p)[1:7],
             col = viridis(3),
             labSize = 3)

```

### 3. Plot counts
####

```{r example_counts, fig.show="hold", out.width="50%", eval=FALSE, include=TRUE, echo=FALSE}

# search genes
anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()

anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()

# plap6
PLAP6 <- "Cre03.g188700"
d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)

resultsNames(dds)
res <- results(dds,contrast = c("genotype","cia5","WT"))
res["Cre03.g188700",]
ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
   # scale_fill_manual(values=cols[c(1,5)]) +
  #  scale_color_manual(values=cols[c(1,5)]) +
   scale_y_continuous(trans = "log2") +
  scale_fill_manual(values = c("grey30","salmon")) + 
  scale_color_manual(values = c("grey30","salmon")) + 
    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
  assign(paste("gcounts_",ig,sep=""),gcounts)

  gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
# ggsave(paste(outdir,"graphs/2023_11_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
#   width = 18,
#   height = 10)

goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "DCRY2", "PHOT1", "UVR8", "HKR1") #"HKR2"

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Get Count data
i <- 1
l <- length(goi_phot)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=anno_phot[i,"gene_id"], intgroup=c("condition","genotype","treatment"), main=anno_phot[i,"id.symbol"],returnData=TRUE)
  d$Gene <- rep(anno_phot[i,"id.symbol"],length(rownames(d)))
  d$sample <- rownames(d)
  # rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "PHOT1", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","treatment","genotype"),col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene, levels = anno_tf$geneSymbol[c(9,6,8,7,15,2,1,4,3,5,13,11,14,10,12)])
levels(all_counts$Gene)

# colours
# group.colors <- rep(anno_colors$treatment, each = 2)
group.colors <- rep(anno_colors$treatment, each = 2)[c(1:6,7,9,11)]
names(group.colors) <- colData(dds)$condition %>% levels()
anno_colors$condition <- group.colors

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts

cols = brewer.pal(n=9,name = 'Paired')

gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition, color=condition)) +
  geom_boxplot(color="black") +
  geom_point(position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=cols) +
  scale_color_manual(values=cols) +
  scale_y_continuous(trans = "log2")
gcounts

gcounts2 <- ggplot(all_counts, aes(x = condition, y = count, fill=Gene)) +
  geom_boxplot(fatten = 1) +
  scale_fill_viridis_d() +
  scale_y_continuous(trans = "log2")
gcounts2

gcounts3 <- ggplot(all_counts[all_counts$Gene=="ROC40",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
  geom_boxplot(color="black") +
  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=anno_colors$treatment) +
  scale_color_manual(values=anno_colors$treatment) +
  scale_y_continuous(trans = "log2") +
  labs(title = "ROC40") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
gcounts3

gcounts4 <- ggplot(all_counts[all_counts$Gene=="ROC114",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
  geom_boxplot(color="black") +
  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
  scale_fill_manual(values=anno_colors$treatment) +
  scale_color_manual(values=anno_colors$treatment) +
  scale_y_continuous(trans = "log2") +
  labs(title = "ROC114") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
gcounts4

```

```{r}
sessionInfo()
```