### 2024-04-21 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
include pCRY figures

---
title: "RNAseq aCRY pCRY data dive"
author: "Kelterborn"
date: "2024-04-19"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    })
    rmarkdown::render(input_file,output_file= 'Readme.md')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

BiocManager::install("ggalt") # , type="source"

# Load System
## -R_libraries
```{r r_libraries}
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(DESeq2)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library(ggpubr)
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)

# library(wget)

```

## -R_folders
```{r r_folders}

# Linux Workstation
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
    ifelse(Sys.info()["sysname"]== "Darwin",
       s <- "/mnt/s",
       s <- "S:"))
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
# list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
# list.files(gitdir) %>% head()

# Macbook
scriptdir <- rstudioapi::getSourceEditorContext()$path
gitdir <- scriptdir %>% dirname() %>% dirname()
dir <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq aCRY"

```

# 1. Load aCRY & pCRY
## - Tximeta files
```{r get_tximeta_files}

# Annotation file
load(file=paste(dir,"anno.RDS", sep="/"))

# Seq 1
tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm"
load(file=tximeta_pcry_file)
gse_pcry <- gse

# Seq 2 (acry + cia5)
load(file=paste(dir,"tximeta.RDS", sep="/"))
gse_acry <- gse[,colData(gse)$experiment=="acry"]
colData(gse_acry) <- colData(gse_acry) %>% droplevels()
colData(gse_acry)$genotype <- colData(gse_acry)$genotype %>% relevel(ref="WT")

# combine
( mcols(gse_acry) == mcols(gse_pcry) ) %>% summary()
# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
names(colData(gse_pcry)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
colData(gse_acry) <- colData(gse_acry)[,c(1:4,7:12)]
colData(gse_pcry)$experiment <- "pcry"
colData(gse_pcry) <- colData(gse_pcry)[,c(1,3,2,4:6,8,12,7,11)]  
# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()

# now same metadata
gse <- cbind(gse_acry,gse_pcry)

colData(gse)$genotype

colData(gse)$treatment <- colData(gse)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
  str_replace(pattern="BL", replacement = "blue") %>%
  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))

colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","pcry_dark","WT_blue","acry_blue","pcry_blue","WT_red","acry_red","pcry_red"))

colData(gse)$experiment

```

## Run Deseq2
```{r dds_acry_pcry}
# design <- ~condition
design <- ~genotype+treatment+genotype:treatment

dds <- DESeqDataSet(gse, design=design)
colData(dds) %>% head() %>% kable()

##########################################\n
# filter all rows with rowsum = 0 ####\n
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

# min counts
# at least 5 counts in 3 samples
keep.sn <- rowSums(counts(dds) >= 5) >= 3
keep.sn %>% summary()
dds <- dds[keep.sn,]

dds <- estimateSizeFactors(dds)
dds
head(assays(dds)[["counts"]])[1:5,1:5]
length(rownames(counts(dds)))

# colData(dds) %>% head() %>% kable()
# add metadata
# colData(dds)$strain <- factor(colData(dds)$strain)

# ## combine technical replicates #####\n
# colData(dds)
# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
# colData(dds)
# dds$runsCollapsed
# colnames(dds)
# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))

###########\n
# run DESeq
dds <- DESeq(dds)

plotDispEsts(dds)
resultsNames(dds)
DESeq2::plotMA(dds)

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$experiment)

head(mcols(dds)$geneSymbol)

###############\n
# save dds #####

# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

```

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
ntd <- normTransform(dds)
rld <- rlog(dds, blind=FALSE) #regularized logarithm
```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
# load(file=paste(data,"rlog.rld", sep="/"))
meanSdPlot(counts(dds, normalized =TRUE))
meanSdPlot(assay(ntd))
meanSdPlot(assay(vsd))
meanSdPlot(assay(rld))

```

### - Check sample distance
```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                    genotype = c("green4","orchid","tan2"))

names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         annotation_col=anno_col,
         annotation_colors = anno_colors,
         show_colnames     = FALSE,
         col=viridis(20),
         cutree_rows = 6,
         cutree_cols = 6,
         fontsize_row = 5)

```

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))
vst_dat <- assay(vst(dds))

### PCA with top 500 genes with highest row variance 
pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE)
# levels(pcaData$genotype)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "treatment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment)

g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "experiment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = anno_colors$genotype)

g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "condition") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  # coord_fixed()
  scale_color_viridis_d(option ="viridis")

((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")

```

###### -- Advanced PCA
```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# calculate PCA (all data)
## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots

# p <- pca(vst_dat, metadata = colData(dds))
# vst <- assay(vst(dds))
p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)

# check different PCAs
p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment

p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype

p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition

(p1+p2) / (p4)

# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

# plot Scree
# find explaining PCs
horn <- parallelPCA(vst_dat)
# horn$n = 2
# warnings()

# horn <- list()
# horn$n <- 2

screeplot(p,
    components = getComponents(p, 1:20),
    vline = c(horn$n, elbow)) +
    geom_label(aes(x = horn$n + 1, y = 30,
      label = paste('Horn',horn$n,sep="="), vjust = -1, size = 8)) +
    geom_label(aes(x = elbow + 1, y = 50,
      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8))

bi <- biplot(p,x="PC3",y="PC1",
    lab = p$metadata$condition,
    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    labSize = 3, legendIconSize = 4.0,
    legendPosition = 'bottom',
    sizeLoadingsNames = 3,
    axisLabSize = 10,
    captionLabSize = 1)

bi <- bi +theme(panel.background = element_rect(fill='transparent'))

pairs <- pairsplot(p,
    components = getComponents(p, c(1:7)),
    triangle = TRUE, trianglelabSize = 6,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),
    plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

layout <- c(
  area(t = 1, l = 1, b = 6, r = 6),
  area(t = 5, l = 1, b = 7, r = 3)
)
pairs + bi + 
  plot_layout(design = layout)

# pairs + inset_element(bi, left = 0, bottom = 0, right = 0.5, top = 0.5)

plotloadings(p,
             components = getComponents(p)[1:7],
             col = viridis(3),
             labSize = 3)

```

# 3. Results

## Plot Counts
```{r}

# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Combined counts table TFs

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

```

# ---------------------------
# COPY & PASTE from pCRY
## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}
outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
setwd(outdir)

dds <- readRDS(paste(outdir,"dds_23design.RDS",sep="/"))
anno <- readRDS(file=paste(outdir,"anno.RDS", sep="/"))

design(dds)
resultsNames(dds)
cbind(resultsNames(dds),c(rep(0,6)))

# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))

# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")

res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))

res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))

# Unterschiede zwischen versch. Belichtungen über beide Genotypen
# D vs. (BL+R)
res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))

# D vs. BL
res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))

# D vs. R
res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))

# WT vs. pCRY
res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))

## Make top lists:

topgenes <- function(res){
a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
a$id.symbol <- anno[rownames(a),"id.symbol"]
a$symbol <- anno[rownames(a),"geneSymbol"]
a <- a[order(a$baseMean, decreasing = T),]
  a$rank.bm <- seq(1:length(rownames(a)))
a <- a[order(a$padj, decreasing = F),]
  a$rank.padj <- seq(1:length(rownames(a)))
a <- a[order(abs(a$log2FoldChange), decreasing = T),]
  a$rank.l2FC <- seq(1:length(rownames(a)))
a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
  a <- a[order(a$rank.sum),]
print(a)
}

top.res1 <- topgenes(res=res1)
summary(res1)
top.res2 <- topgenes(res=res2)
summary(res2)
top.res3 <- topgenes(res=res3)
summary(res3)
top.res4 <- topgenes(res=res4)
summary(res4)
top.res5 <- topgenes(res=res5)
summary(res5)
top.res6 <- topgenes(res=res6)
summary(res6)
top.res7 <- topgenes(res=res7)
summary(res7)

top.all <- rbind(top.res1,
         top.res2,
         top.res3,
         top.res4,
         top.res5,
         top.res6,
         top.res7)
top.all <- top.all[order(top.all$rank.sum),]
head(top.all, n=185) %>% rownames() %>% unique() %>% length()
top.all <- top.all[unique(rownames(top.all)),]

# top.all <- head(top.all, n=100)

## Shrinkage
res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
res7_shrink <- lfcShrink(dds, res=res7, type="ashr")

```

### combine L2FC with padj
```{r}

res3_shrink
subset(mcols(dds), id.symbol =="SELU1") %>% rownames()
res3_shrink["Cre06.g263550",]

# DataFrame with 1 row and 5 columns
#                baseMean log2FoldChange     lfcSE      pvalue        padj
#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
# Cre06.g263550   3566.14        2.90397  0.167321 1.72033e-71 1.97322e-68

plot(res3_shrink$log2FoldChange,-log(res3_shrink$padj))
res3_shrink$logpadj <- -log(res3_shrink$padj, base=10)

-log(1.38379e-06,base = 10)

x=-2.5
60/(3^2)

y=-8*x^2 + 50
y

res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < res3_shrink$logpadj,TRUE,FALSE)
summary(res3_shrink$ishs)
subset(res3_shrink,ishs==TRUE & logpadj< 50 & abs(log2FoldChange)<2.5)
subset(res3_shrink,ishs==TRUE & logpadj< 50)

# for log2 -/+ 5 and padj 100 
res1_shrink$ishs <- ifelse((-20/3*(res1_shrink$log2FoldChange^2) + 60) < -log(res1_shrink$padj, base=10),TRUE,FALSE)
res2_shrink$ishs <- ifelse((-20/3*(res2_shrink$log2FoldChange^2) + 60) < -log(res2_shrink$padj, base=10),TRUE,FALSE)
res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < -log(res3_shrink$padj, base=10),TRUE,FALSE)
res4_shrink$ishs <- ifelse((-20/3*(res4_shrink$log2FoldChange^2) + 60) < -log(res4_shrink$padj, base=10),TRUE,FALSE)
res5_shrink$ishs <- ifelse((-20/3*(res5_shrink$log2FoldChange^2) + 60) < -log(res5_shrink$padj, base=10),TRUE,FALSE)
res6_shrink$ishs <- ifelse((-20/3*(res6_shrink$log2FoldChange^2) + 60) < -log(res6_shrink$padj, base=10),TRUE,FALSE)
res7_shrink$ishs <- ifelse((-20/3*(res7_shrink$log2FoldChange^2) + 60) < -log(res7_shrink$padj, base=10),TRUE,FALSE)

```

```{r}
outdir
figures <- paste(outdir,"figures",sep="/")
dir.create(figures)

```

## Colours
```{r}
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
```

## PCA (Panel B)
```{r}
library("PCAtools")
library(ggalt)

vsd <- vst(dds, blind=FALSE)
length(vsd)

dubs <- duplicated(mcols(vsd)$id.symbol)
length(dubs)

unique <- unique(mcols(vsd)$id.symbol)
length(unique)

vsd2 <- vsd[!duplicated(mcols(vsd)$id.symbol)]
rownames(vsd2) <- mcols(vsd2)$id.symbol

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
names(group.colors) <- levels(colData(dds)$condition)
group.colors

plotPCA(vsd2, intgroup="condition")
vst <- assay(vsd2)
p <- pca(vst, metadata = colData(dds), removeVar = 0.1)
screeplot(p, axisLabSize = 18, titleLabSize = 22)
biplot(p, colby = "condition")
biplot(p, showLoadings = TRUE,labSize = 5, pointSize = 5, sizeLoadingsNames = 5)

biplot(p,x= "PC1", y="PC3",
       lab = p$metadata$sample,
    colby = 'condition', colkey = group.colors,
    encircle = TRUE,
    encircleFill = TRUE,
    legendPosition = 'right',
    labSize = 4)
ggsave(paste(figures,"2023_12_PCA1-3.png",sep="/"),
  width = 6,
  height = 6)

biplot(p,x= "PC3", y="PC4",
       lab = p$metadata$sample,
    colby = 'condition', colkey = group.colors,
    encircle = TRUE,
    encircleFill = TRUE,
    legendPosition = 'right')
ggsave(paste(figures,"2023_12_PCA3-4.pdf",sep="/"),
  width = 12,
  height = 12)

pairsplot(p,colby = 'condition', colkey = group.colors)
ggsave(paste(figures,"2023_12_pairsplot.pdf",sep="/"),
  width = 12,
  height = 12)

plotloadings(p,
    rangeRetain = 0.01,
    labSize = 2.0,
    title = 'Loadings plot',
    subtitle = 'PC1, PC2, PC3, PC4, PC5',
    caption = 'Top 1% variables',
    drawConnectors = TRUE)
ggsave(paste(figures,"2023_12_loadings.pdf",sep="/"),
  width = 12,
  height = 12)

# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

anno["Cre01.g016600",]

 biplot(p,
    lab = p$metadata$condition,
    colby = 'condition', colkey = group.colors,
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    legendLabSize = 16, legendIconSize = 8.0,
    legendPosition = 'right')

pairsplot(p,
    components = getComponents(p, c(1:6)),
    triangle = TRUE, trianglelabSize = 12,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = group.colors,
    title = 'Pairs plot', plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

```

## Counts
```{r}

# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Combined counts table TFs

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

```

## Volcanos
### dark
```{r}

# dark
res <- res1_shrink
n <- 'WT vs. pCRY in dark'
nf <- 'figures/Volcano_dark.png'
l <- nrow(top.res1)
pcol <- "black"
lcol <- colours[1+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evd <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-60),
    FCcutoff = 3,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evd

```

### blue
```{r}

# blue
res <- res2_shrink
n <- 'WT vs. pCRY in blue'
nf <- 'figures/Volcano_blue.png'
l <- nrow(top.res2)
pcol <- colours[3]
lcol <- colours[3+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 17,
                ifelse(res$padj < 1e-100,17,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evb <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evb

```

### red
```{r}

# red
res <- res3_shrink
n <- 'WT vs. pCRY in red'
nf <- 'figures/Volcano_r.png'
l <- nrow(top.res3)
pcol <- colours[5]
lcol <- colours[5+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evr <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evr

#ggsave(file="figures/Vulcano_rtest.pdf",
#       width = 12,
#       height = 12)

```

### all
```{r}

ga <- ggarrange(evb,
                evd+rremove("ylab"),
                evr+rremove("ylab"),
                ncol = 3, nrow = 1,common.legend = TRUE, legend = "bottom",
                widths = c(1.05,1,1))

ggsave(ga, file="figures/Vulcano_all.pdf",
       width = 12,
       height = 8)

# + ggplot2::coord_cartesian(xlim=c(-10, 10))
# ggplot2::coord_cartesian(xlim=c(-5, 5)

# with ggplot
ggplot(as.data.frame(resLFC_WT_BL_vs.pcry_BL),aes(log2FoldChange,-log(padj))) +
  geom_point() +
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rownames(top_WT_BL_vs.pcry_BL),]), colour = "blue") + 
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), colour = "red") +
  geom_text_repel(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), aes(label=anno[rocs,"id.symbol"]),colour = "red",hjust=-0.5, vjust=-1) +
  xlim(-5,5)
getwd()
ggsave("graphs/Vulcano ROCs-2.pdf",
       width = 10,
       height = 6)

## enhanced vulcano
library(airway)
  library(magrittr)

 library('DESeq2')

  dds <- DESeqDataSet(airway, design = ~ cell + dex)
  dds <- DESeq(dds, betaPrior=FALSE)
  res <- results(dds,
    contrast = c('dex','trt','untrt'))
  res <- lfcShrink(dds,
    contrast = c('dex','trt','untrt'), res=res, type = 'normal')
 EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    y = 'pvalue')

```

## Heatmap

### plot heatmap
```{r}
library("pheatmap")
ntd <- normTransform(dds)
top.all.hm <- head(top.all, n=100)
df <- assay(ntd)[rownames(top.all.hm),]
rownames(df) <- mcols(dds)[rownames(df),"id.symbol"]

# anno_col <- as.data.frame(colData(dds)[,c("condition","treatment","genotype")])
anno_col <- data.frame(condition = dds$condition, treatment = dds$treatment)
rownames(anno_col) <- rownames(colData(dds))

# column order
anno_col <- arrange(anno_col, match(anno_col$condition, levels(anno_col$condition)))
levels(anno_col$condition)
df <- df[,rownames(anno_col)]

annot_colors=list(condition=c(Cancer="#F0978D",Healthy="#63D7DE"))
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)

mycolors2 <- list(genotype = c("#525252","#969696"),
     treatment = c("#525252","#1F78B4","#E31A1C"),
     condition = colours)

names(mycolors2$genotype) <- levels(anno_col$genotype)
names(mycolors2$treatment) <- levels(anno_col$treatment)
names(mycolors2$condition) <- levels(anno_col$condition)
mycolors2

xx <- pheatmap(df, cluster_rows=TRUE, show_rownames=TRUE,
         cluster_cols=FALSE,
         annotation_col=anno_col,
         color=colorRampPalette(c("navy", "white", "red"))(50),
         annotation_colors=mycolors2
         )

ggsave("figures/2023_12_Heatmap_top.png",plot=xx,
       width = 5,
       height = 15)

```

### Complex Heatmap
```{r}

df
genes <- anno_tf

# collapse replicates
colData(dds)
dds_heat <- collapseReplicates(dds, dds$condition,dds$sample)
dds_heat$sample <- factor(dds_heat$sample, levels=c("WT-D1","pcry-D1","WT-BL1","pcry-BL1","WT-R1","pcry-R1"))
colData(dds_heat)
dds_heat$runsCollapsed
colnames(dds_heat)
matchFirstLevel <- dds_heat$sample == levels(dds_heat$sample)[1]
stopifnot(all(rowSums(counts(dds_heat[,matchFirstLevel])) == counts(dds_heat[,1])))

mat <- counts(dds_heat)
length(rownames(counts(dds_heat)))
length(rownames(df))
rownames(mat) <- mcols(dds_heat)$id.symbol
length(rownames(mat))

mat <- mat[df,]

position <- mat
l <- length(rownames(mat))
pos <- data.frame(genes = rownames(mat),pos = 1:l)
rownames(pos) <- rownames(mat)
pos <- pos[genes,2]

# mat <- cbind(res,counts(dds_heat))
# mat <- subset(mat,!SYMBOL2=="" & SYMBOL2 %in% df$SYMBOL2)
# rownames(mat) <- mat$SYMBOL2
# mat <- mat[,c(12:15)]
heat <- t(scale(t(mat)))
Heatmap(heat)

summary(heat)
myCol <- colorRampPalette(c('brown3', 'yellow2', 'darkolivegreen3'))(100)
myBreaks <- seq(-1.5, 1.5, length.out = 100)

ann <- data.frame(condition = colData(dds_heat)[,"condition"],
  stringsAsFactors = FALSE)

colAnn <- HeatmapAnnotation(
  df = ann,
  which = 'col', # 'col' (samples) or 'row' (gene) annotation?
  na_col = 'white', # default colour for any NA values in the annotation data-frame, 'ann'
  # col = colours,
  annotation_height = 0.6,
  annotation_width = unit(1, 'cm'),
  gap = unit(1, 'mm'))

boxplotCol <- HeatmapAnnotation(
  boxplot = anno_boxplot(
    heat,
    border = FALSE,
    gp = gpar(fill = '#CCCCCC'),
    pch = '.',
    size = unit(2, 'mm'),
    axis = TRUE,
    axis_param = list(
      gp = gpar(fontsize = 12),
      side = 'left')),
  annotation_width = unit(c(2.0), 'cm'),
  which = 'col')

boxplotRow <- HeatmapAnnotation(
  boxplot = row_anno_boxplot(
    heat,
    border = FALSE,
    gp = gpar(fill = '#CCCCCC'),
    pch = '.',
    size = unit(2, 'mm'),
    axis = TRUE,
    axis_param = list(
      gp = gpar(fontsize = 12),
      side = 'top')),
  annotation_width = unit(c(2.0), 'cm'),
  which = 'row')

genelabels <- rowAnnotation(
  Genes = anno_mark(
    at = pos,
    labels = genes,
    labels_gp = gpar(fontsize = 15, fontface = 'bold'),
    padding = 0.75),
  width = unit(2.0, 'cm') +

    max_text_width(
      rownames(heat)[seq(1, nrow(heat), 40)],
      gp = gpar(fontsize = 10,  fontface = 'bold')))

# Clusters
# pamClusters <- cluster::pam(heat, k = 4) # pre-select k = 4 centers
pamClusters <- cluster::pam(heat, k= 8) # pre-select k = 4 centers

pamClusters$clustering <- paste0('Cluster ', pamClusters$clustering)
pamClusters$clustering <- factor(pamClusters$clustering,
                                 levels = c('Cluster 1', 'Cluster 2', 'Cluster 3', 'Cluster 4','Cluster 5', 'Cluster 6'))

hmap <- Heatmap(heat,

                # split the genes / rows according to the PAM clusters
                split = pamClusters$clustering,
                cluster_row_slices = TRUE,

                name = 'Gene\nZ-\nscore',

                col = colorRamp2(myBreaks, myCol),

                # parameters for the colour-bar that represents gradient of expression
                heatmap_legend_param = list(
                  color_bar = 'continuous',
                  legend_direction = 'vertical',
                  legend_width = unit(8, 'cm'),
                  legend_height = unit(5.0, 'cm'),
                  title_position = 'topcenter',
                  title_gp=gpar(fontsize = 12, fontface = 'bold'),
                  labels_gp=gpar(fontsize = 12, fontface = 'bold')),

                # row (gene) parameters
                cluster_rows = TRUE,
                show_row_dend = TRUE,
                row_title = 'Statistically significant genes',
                row_title_side = 'left',
                row_title_gp = gpar(fontsize = 12,  fontface = 'bold'),
                row_title_rot = 90,
                show_row_names = FALSE,
                row_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                row_names_side = 'left',
                row_dend_width = unit(25,'mm'),

                # column (sample) parameters
                cluster_columns = TRUE,
                show_column_dend = TRUE,
                column_title = '',
                column_title_side = 'bottom',
                column_title_gp = gpar(fontsize = 12, fontface = 'bold'),
                column_title_rot = 0,
                show_column_names = TRUE,
                column_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                column_names_max_height = unit(10, 'cm'),
                column_dend_height = unit(25,'mm'),

                # cluster methods for rows and columns
                clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
                clustering_method_columns = 'ward.D2',
                clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
                clustering_method_rows = 'ward.D2',

                # specify top and bottom annotations
                top_annotation = colAnn)
                # bottom_annotation = boxplotCol)

pdf("Heatmap2.pdf", height = 18, width = 6)
draw(hmap + genelabels,
     heatmap_legend_side = 'left',
     annotation_legend_side = 'right',
     row_sub_title_side = 'left')
dev.off()

```

# Export

```{r}

library(writexl)

# combine results

res <- res_WT_BL_vs.pcry_BL
res1 <- res_WT_D_vs.pcry_D
res2 <- res_WT_BL_vs.pcry_BL
res3 <- res_WT_R_vs.pcry_R
res4 <- res_WT_D_vs.WT_BL
res5 <- res_WT_D_vs.WT_R

colnames(anno)

res_exp <- data.frame(
  # general
  "gene_id" = rownames(dds),
  "gene_name" = anno[rownames(dds),"geneSymbol"],
  "Alias" = anno[rownames(dds),"prev.symbols"],
  "id.symbol" = anno[rownames(dds),"id.symbol"],
 "Description" = anno[rownames(dds),"Description"],
 "Comments" = anno[rownames(dds),"Comments"],
 "TargetP" = anno[rownames(dds),"TargetP"],
 "Predalgo" = anno[rownames(dds),"Predalgo"],
 "baseMean" = res$baseMean,
 # results 1
 "L2FC.WT_D_vs.pcry_D" = res1$log2FoldChange,
 "pvalue.WT_D_vs.pcry_D" = res1$pvalue,
 "padj.WT_D_vs.pcry_D" = res1$padj,
  # results 2
 "L2FC.WT_BL_vs.pcry_BL" = res2$log2FoldChange,
 "pvalue.WT_BL_vs.pcry_BL" = res2$pvalue,
 "padj.WT_BL_vs.pcry_BL" = res2$padj,
  # results 3
 "L2FC.WT_R_vs.pcry_R" = res3$log2FoldChange,
 "pvalue.WT_R_vs.pcry_R" = res3$pvalue,
 "padj.WT_R_vs.pcry_R" = res3$padj,
  # results 4
 "L2FC.WT_D_vs.WT_BL" = res4$log2FoldChange,
 "pvalue.WT_D_vs.WT_BL" = res4$pvalue,
 "padj.WT_D_vs.WT_BL" = res4$padj,
  # results 5
 "L2FC.WT_D_vs.WT_R" = res5$log2FoldChange,
 "pvalue.WT_D_vs.WT_R" = res5$pvalue,
 "padj.WT_D_vs.WT_R" = res5$padj,
  counts(dds, normalized=TRUE))
res_exp

outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"

write_xlsx(data.frame(res_exp),
           paste(outdir,"2023_08 P3044 results.xlsx",sep="/"))

# Counts only
counts <- round(counts(dds, normalized=TRUE))
colnames(counts)
colData(dds)$sample.name <- paste(colData(dds)$condition,colData(dds)$repetition,sep="_")
colnames(counts) == colData(dds)$sample.id
colnames(counts) <- colData(dds)$sample.name

write.csv(counts, file.path(outdir,"Counts.csv"))

Exp_design <- data.frame(names = colData(dds)$sample.name,
                         genotype = colData(dds)$genotype,
                         treatment = colData(dds)$treatment,
                         condition = colData(dds)$condition) %>% t()

Exp_design <- matrix(colData(dds)$sample.name,
                     colData(dds)$genotype,
                     colData(dds)$treatment,
                    colData(dds)$condition)

rownames(Exp_design)[1] <- {1}

write.table(Exp_design,file.path(outdir,"Exp_design.csv"), sep=",",  col.names=FALSE)

```

```{r}
sessionInfo()
```