### 2024-05-17 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
Vulcanos

---
title: "RNAseq aCRY pCRY data dive (separate)"
author: "Kelterborn"
date: "2024-04-19"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'Readme.md')
        })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

BiocManager::install()

# Load System
## -R_libraries
```{r r_libraries, echo=FALSE}
library(patchwork)
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(DESeq2)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)

library(ggpubr)
library(vsn)
library(stringr)
library(R.utils)
library(RColorBrewer)

# library(wget)

```

## -R_folders
```{r r_folders, include = FALSE}

# Linux Workstation
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
    ifelse(Sys.info()["sysname"]== "Darwin",
       s <- "/mnt/s",
       s <- "S:"))
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
# list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
# list.files(gitdir) %>% head()

# Macbook
scriptdir <- rstudioapi::getSourceEditorContext()$path
gitdir <- scriptdir %>% dirname() %>% dirname()

if(Sys.info()["sysname"]== "Linux"){
  s <- "/mnt/s"
  dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")} else {
    if(Sys.info()["sysname"]== "Darwin"){
       s <- "/mnt/s"
       dir <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq aCRY"} else {
       s <- "S:" 
       dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")}}

```

## -R functions
```{r f_get_results}
getresults_SK <- function(contrast,nres){
r1 <- results(dds, contrast = contrast)
r1$symbol <- mcols(dds)$symbol
assign(paste("res",nres,sep="."),r1)
r1 <- list(r1)
names(r1) <- nres
r1
}
```

```{r f_topgenes}
topgenes_f <- function(res,p=0.5,bM=20,l2FC=1){
a <- subset(res, padj < p & baseMean > bM & abs(log2FoldChange) > l2FC)
a <- a[order(a$baseMean, decreasing = T),]
  a$rank.bm <- seq(1:length(rownames(a)))
a <- a[order(a$padj, decreasing = F),]
  a$rank.padj <- seq(1:length(rownames(a)))
a <- a[order(abs(a$log2FoldChange), decreasing = T),]
  a$rank.l2FC <- seq(1:length(rownames(a)))
a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
  a <- a[order(a$rank.sum),]
a
}
```

```{r f_vulcano}

Vulcano_SK <- function(res,
                       n=NULL,
                       list1=NULL,
                       list2=NULL,
                       list1.col=NULL,
                       list2.col=NULL,
                       list1.n=NULL,
                       list2.n=NULL,
                       xlim=NULL,
                       ylim=NULL,
                       ...)
{
# Name
n <- names(res)
# unlist results
if(is.list(res)){
  res <- res[[1]]
}
# define list 1
if(is.null(list1) ) {
  list1 <- topgenes_f(res,p=0.05,bM = 10,l2FC = 1)
}
# define list 2
if(is.null(list2) ) {
  list2 <- topgenes_f(res,p=0.005,bM = 100,l2FC = 2)
}
# list 1 color
if(is.null(list1.col) ) {
  list1.col <- "royalblue1"
}
# list 2 color
if(is.null(list2.col) ) {
  list2.col <- "royalblue4"
}
# name list 1
if(is.null(list1.n) ) {
  list1.n <- "list1"
}
# name list 2
if(is.null(list2.n) ) {
  list2.n <- "list2"
}

# xlim
if(is.null(xlim) ) {
  xlim <- abs(c(min(res$log2FoldChange, na.rm=TRUE),
              max(res$log2FoldChange, na.rm=TRUE))) %>% max()
}

# ylim
if(is.null(ylim) ) {
  ylim <- max(-log10(res$padj), na.rm = TRUE) + 5
}

# colors
pcol <- "black"
lcol <- colours[1+1]

# add symbol
res$symbol <- mcols(dds_list[[i]])$symbol

# define list groups
res$isl1 <- rownames(res) %in% rownames(list1)
res$isl2 <- rownames(res) %in% rownames(list2)

# order results (list 1 on top)
res <- res[order(abs(res$log2FoldChange), decreasing = T),]
res <- res[order(res$isl1, decreasing = F),]
res <- res[order(res$isl2, decreasing = F),]

# change outlier shape
shape <- 1
if(!is.null(xlim)){
shape <- ifelse(abs(res$log2FoldChange) > xlim, 6,16)
}
if(!is.null(ylim)){
shape[res$padj < 10^-ylim] <- 6
}
# name shape
names(shape)[shape == 6] <- 'outlier'
names(shape)[shape == 16] <- 'in range'
# move outlier to border
res$log2FoldChange[res$log2FoldChange > xlim] <- xlim
res$log2FoldChange[res$log2FoldChange < -xlim] <- -xlim
res$padj[res$padj < 10^-ylim] <- 10^-ylim

# name colors
keyvals <- ifelse(
    res$isl2 == TRUE, list2.col,
    ifelse(
    res$isl1 == TRUE, list1.col,
    'grey') )

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == list1.col] <- list1.n
  names(keyvals)[keyvals == list2.col] <- list2.n
  names(keyvals)[keyvals == 'grey'] <- 'other'

ev_f <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = res$symbol,
    selectLab  = res$symbol[which(names(keyvals) == list2.n)],
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = FALSE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 100,
    colCustom = keyvals,
    xlim = c(-xlim,xlim),
    ylim = c(0, ylim),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("list1:",nrow(list1),", list2: ",nrow(list2)),
    # sub = "SVF",
    # pCutoff = 10^(-60),
    # FCcutoff = 1,
    shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2','p-adj < 0.05',
                    'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
ev_f

}

```

# 1. Load aCRY & pCRY
## - Tximeta files
```{r get_tximeta_files, include=FALSE}

# Annotation file
load(file=paste(dir,"anno.RDS", sep="/"))

# Seq 1 (pcry)
ifelse(Sys.info()["sysname"]== "Linux",
       tximeta_pcry_file <- dir %>% dirname() %>% paste(.,"P3044/tximeta.txm",sep="/"),
    ifelse(Sys.info()["sysname"]== "Darwin",
       tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm",
       tximeta_pcry_file <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044/tximeta.txm"))

load(file=tximeta_pcry_file)

colData(gse)$treatment <- colData(gse)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
  str_replace(pattern="BL", replacement = "blue") %>%
  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))

colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","pcry_dark","WT_blue","pcry_blue","WT_red","pcry_red"))

names(colData(gse)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
colData(gse)$experiment <- "pcry"
gse_pcry <- gse

# Seq 2 (acry + cia5)
load(file=paste(dir,"tximeta.RDS", sep="/"))
gse_acry <- gse[,colData(gse)$experiment=="acry"]
colData(gse_acry) <- colData(gse_acry) %>% droplevels()
colData(gse_acry)$genotype <- colData(gse_acry)$genotype %>% relevel(ref="WT")

colData(gse_acry)$treatment <- colData(gse_acry)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
  str_replace(pattern="BL", replacement = "blue") %>%
  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))

colData(gse_acry)$condition <- paste(colData(gse_acry)$genotype,colData(gse_acry)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red"))

```

## Run Deseq2
```{r dds_acry_pcry, echo=FALSE}

# design <- ~condition
exp <- c("acry","pcry")
gse_list <- list(gse_acry,gse_pcry)
names(gse_list) <- exp
dds_list <- list()
i <- 1
for (e in names(gse_list)){
 gse <- gse_list[[e]]
 design <- ~genotype+treatment+genotype:treatment
dds <- DESeqDataSet(gse, design=design)
colData(dds) %>% head() %>% kable()

##########################################\n
# filter all rows with rowsum = 0 ####\n
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10), main = colData(dds)$experiment[1]) 
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

# min counts
# at least 5 counts in 3 samples
keep.sn <- rowSums(counts(dds) >= 5) >= 3
keep.sn %>% summary()
dds <- dds[keep.sn,]

dds <- estimateSizeFactors(dds)
dds
head(assays(dds)[["counts"]])[1:5,1:5]
length(rownames(counts(dds)))

# remove outlier
outliner <- c("WTdark3")
colData(dds)[outliner,]
keep <- colnames(dds) %>% .[!str_detect(.,outliner)]
dds <- dds[,keep]

###########\n
# run DESeq
dds <- DESeq(dds)

plotDispEsts(dds, main = colData(dds)$experiment[1])
resultsNames(dds)
DESeq2::plotMA(dds, main=colData(dds)$experiment[1])

ROC40 <- "Cre06.g275350"
plotCounts(dds, gene = ROC40, intgroup = "condition", col=colData(dds)$treatment, main = paste("ROC40 in Δ",colData(dds)$experiment[1],sep = ""))

assign(paste("dds",e,sep="_"),dds)
dds_list[[e]] <- dds
}

###############\n
# save dds #####

# save(dds_list, file = paste(dir,"dds_list.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
# load(paste(dir,"dds.RDS",sep="/"))
# dds

```

# 2. Pre-Analysis
### - Data transformations
```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}

vsd_list <- lapply(dds_list,vst,blind=FALSE)
ntd_list <- lapply(dds_list,normTransform)
rld_list <- lapply(dds_list,rlog, blind=FALSE)
# save(rld_list, file = paste(dir,"rlog_list.RDS",sep="/"), compress = FALSE)

```

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="50%"}
# load(file = paste(dir,"rlog_list.RDS",sep="/"))
msdp1 <- lapply(lapply(dds_list,assay,normalized =TRUE),meanSdPlot)
msdp2 <- lapply(lapply(vsd_list,assay,normalized =TRUE),meanSdPlot)
msdp3 <- lapply(lapply(ntd_list,assay,normalized =TRUE),meanSdPlot)
msdp4 <- lapply(lapply(rld_list,assay,normalized =TRUE),meanSdPlot)

```

### - Check sample distance
```{r pre_sample_dist, fig.height=4, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
vsd <- vsd_list[[1]]
for (vsd in vsd_list){
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- vsd$clientName
# colnames(sampleDistMatrix) <- NULL
colData(vsd)$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype")])
anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
                    genotype = c("green4","tan2"))
names(anno_colors$treatment) <- levels(anno_col$treatment)
names(anno_colors$genotype) <- levels(anno_col$genotype)

pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         annotation_col=anno_col,
         annotation_colors = anno_colors,
         show_colnames     = FALSE,
         col=viridis(20),
         cutree_rows = 6,
         cutree_cols = 6,
         fontsize_row = 5) %>% print()
}
# -> how to handel batch effects: https://www.biostars.org/p/266507/

```

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=5, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))

for (vsd in vsd_list){

### PCA with top 500 genes with highest row variance 
pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE)
# levels(pcaData$genotype)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "treatment") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment)

g2 <- ggplot(pcaData, aes(PC1, PC2, color=genotype, shape=genotype)) +
  geom_point(size=5, alpha=0.7) +
  labs(title = "genotype") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = anno_colors$genotype)

g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype, label=clientName)) +
  geom_point(size=5, alpha=0.7) +
  geom_text_repel() +
  labs(title = "condition") +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  # coord_fixed()
  scale_color_viridis_d(option ="viridis")

gp <- ((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect") & plot_annotation(title = colData(vsd)$experiment[1])
gp %>% print()
}

```

###### -- Advanced PCA
```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}

for (dds in dds_list) {
vst_dat <- assay(vst(dds))
p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)

# check different PCAs
p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 8,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment

p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 8,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype

p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 8,trianglelabSize = 6,
                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition

pairs <- (p1+p2) / (p4) + plot_annotation(title = colData(dds)$experiment[1])
pairs %>% print()
# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

# plot Scree
# find explaining PCs
horn <- parallelPCA(vst_dat)
# horn$n = 4 #aCRY
# warnings()

# horn <- list()
# horn$n <- 4

s <- screeplot(p,
    components = getComponents(p, 1:20),
    vline = c(horn$n, elbow)) +
    geom_label(aes(x = horn$n + 1, y = 30,
      label = paste('Horn',horn$n,sep="="), vjust = -1, size = 8)) +
    geom_label(aes(x = elbow + 1, y = 50,
      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8)) + plot_annotation(title = colData(dds)$experiment[1])

s %>% print()

bi <- biplot(p,x="PC3",y="PC1",
    lab = p$metadata$clientName,
    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    labSize = 3, legendIconSize = 4.0,
    legendPosition = 'bottom',
    sizeLoadingsNames = 3,
    axisLabSize = 10,
    captionLabSize = 1)

bi <- bi +theme(panel.background = element_rect(fill='transparent'))

pairs <- pairsplot(p,
    components = getComponents(p, c(1:7)),
    triangle = TRUE, trianglelabSize = 6,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),
    plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

layout <- c(
  area(t = 1, l = 1, b = 6, r = 6),
  area(t = 5, l = 1, b = 7, r = 3)
)
pairs_bi <- pairs + bi + 
  plot_layout(design = layout) & plot_annotation(title = colData(dds)$experiment[1])
pairs_bi %>% print()

# pairs + inset_element(bi, left = 0, bottom = 0, right = 0.5, top = 0.5)

load <- plotloadings(p,
                     axisLabSize = 8,
             components = getComponents(p)[1:7],
             col = viridis(3),
             labSize = 3) + 
  plot_annotation(title = colData(dds)$experiment[1]) &
  theme(legend.position = 'none', 
        axis.title=element_text(size=8),
        axis.text=element_text(size=6))
load %>% print()
}

```

# 3. Results
## Colors
```{r colors, echo=FALSE}
# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
# group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")

anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                    genotype = c("green4","tan2","orchid"))

group.colours2 <- c("#00B500","#007E56","#805B00","#FF00AB","#8024AB","#FF0056","tan2","brown","orangered")
```

## Plot Counts
```{r plot_counts, echo=FALSE, fig.width=12}
dds <- dds_list[[1]]
for (dds in dds_list){
# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts_tf1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_tf1 %>% print()

# ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
# width = 12,
# height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Combined counts table TFs

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# Plot
max_val <- 1.0*max(all_counts$count)
gcounts_phot1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_phot1 %>% print()

# print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))

# single gene
    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)

gcounts <- ggplot(d, aes(x = treatment, y = count, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    labs(title = paste("ROC40"," in Δ",colData(dds)$experiment[1],sep = ""))
gcounts %>% print()

}

# ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
#   width = 12,
#   height = 8)

```

## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}

design(dds)
resultsNames(dds)
cbind(resultsNames(dds),c(rep(0,resultsNames(dds) %>% length())))

#      [,1]                         [,2]
# [1,] "Intercept"                  "0" 
# [2,] "genotype_pcry_vs_WT"        "0" 
# [3,] "treatment_blue_vs_dark"     "0" 
# [4,] "treatment_red_vs_dark"      "0" 
# [5,] "genotypepcry.treatmentblue" "0" 
# [6,] "genotypepcry.treatmentred"  "0"

# assign(paste("res",n,sep="_"),list())

res_list <- list()
for (i in 1:2) {
dds <- dds_list[[i]]

nlist <- names(dds_list[i])
c <- resultsNames(dds)
print(nlist)
res_l <- list()

# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))

# Belichtungseffect für jeden Genotyp
nres <- "WT_BL.vs.D"
contrast <- list(c("treatment_blue_vs_dark"))
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- "WT_R.vs.D"
contrast <- list(c("treatment_red_vs_dark"))
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- paste(nlist,"_BL.vs.D",sep="")
contrast <- list(c("treatment_blue_vs_dark",c[5])) # [5,] "genotypepcry.treatmentblue"
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- paste(nlist,"_R.vs.D",sep="")
contrast <- list(c("treatment_red_vs_dark",c[6])) # [6,] "genotypepcry.treatmentred"  
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

# Unterschiede zwischen WT und ko unter gleicher Belichtung
nres <- paste(nlist,"D.vs.WT_D",sep="_")
contrast <- list(c[2])                # [2,] "genotype_pcry_vs_WT"
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- paste(nlist,"BL.vs.WT_BL",sep="_")
contrast <- list(c[2],c[5])           # [5,] "genotypepcry.treatmentblue"
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- paste(nlist,"R.vs.WT_R",sep="_")
contrast <- list(c[2],c[6])           # [6,] "genotypepcry.treatmentred"  
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

# Unterschiede im Belichtungseffekt
nres <- paste(nlist,"BLvD.vs.WT_BLvD",sep="_")
contrast <- c(0,0,0,0,1,0)
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- paste(nlist,"RvD.vs.WT_RvD",sep="_")
contrast <- c(0,0,0,0,0,1)
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

## Exotische Kontraste
# Unterschiede zwischen versch. Belichtungen über beide Genotypen
nres <- "BL+R.vs.D"
contrast <- c(0,0,2/6,2/6,1/3,1/3)
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- "BL.vs.D"
contrast <- c(0,0,1,0,0.5,0)
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

nres <- "R.vs.D"
contrast <- c(0,0,0,1,0,0.5)
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

# WT vs. ko
nres <- paste(nlist,".vs.WT",sep="")
contrast <- c(0,1,0,0,1/3,1/3)
res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)

print(length(res_l))
print(length(res_list))
print(names(res_list))

res_list[[nlist]] <- res_l
print(length(res_list))
print(names(res_list))

    }

## Make deg & top lists:
deg_list <- list()
top_list <- list()

for (i in 1:2) {
ko <- names(res_list[i])

deg_genes_l <- list()
deg_genes_l <- lapply(res_list[[i]],topgenes_f) %>%  lapply(.,rownames) 
names(deg_genes_l) <- paste("deg",names(res_list[[i]]),sep="_")
deg_list[[ko]] <- deg_genes_l

top_genes_l <- list()
top_genes_l <- lapply(res_list[[i]],topgenes_f,p=0.01, bM=100, l2FC=2) %>%  lapply(.,rownames) 
names(top_genes_l) <- paste("top",names(res_list[[i]]),sep="_")
top_list[[ko]] <- top_genes_l

}

top_genes_l <- lapply(res_list,topgenes_f,p=0.01, bM=100, l2FC=2) %>%  lapply(.,rownames)

## Make shrinkage
res_ashr_list <- list()
for (i in 1:2) {
ko <- names(res_list[i])

res_ashr_l <- list()
l <- length(res_list[[i]])
for (j in 1:l){
res_ashr_l[[names(res_list[[i]])[j]]] <- lfcShrink(res=res_list[[i]][[j]], dds=dds_list[[i]], type="ashr")
}
res_ashr_list[[ko]] <- res_ashr_l
}

```

## Volcanos
### dark
```{r, eval=TRUE}

i <- 2

Vulcano_SK(res_ashr_list[[i]]["pcry_BL.vs.WT_BL"],xlim=5,ylim=50)

# dark
res <- res1_shrink
n <- 'WT vs. pCRY in dark'
nf <- 'figures/Volcano_dark.png'
l <- nrow(top.res1)
pcol <- "black"
lcol <- colours[1+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evd <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-60),
    FCcutoff = 3,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evd

```

### blue
```{r eval=FALSE}

# blue
res <- res2_shrink
n <- 'WT vs. pCRY in blue'
nf <- 'figures/Volcano_blue.png'
l <- nrow(top.res2)
pcol <- colours[3]
lcol <- colours[3+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 17,
                ifelse(res$padj < 1e-100,17,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evb <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evb

```

### red
```{r, eval=FALSE}

# red
res <- res3_shrink
n <- 'WT vs. pCRY in red'
nf <- 'figures/Volcano_r.png'
l <- nrow(top.res3)
pcol <- colours[5]
lcol <- colours[5+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evr <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evr

#ggsave(file="figures/Vulcano_rtest.pdf",
#       width = 12,
#       height = 12)

```

### all
```{r, eval=FALSE}

ga <- ggarrange(evb,
                evd+rremove("ylab"),
                evr+rremove("ylab"),
                ncol = 3, nrow = 1,common.legend = TRUE, legend = "bottom",
                widths = c(1.05,1,1))

ggsave(ga, file="figures/Vulcano_all.pdf",
       width = 12,
       height = 8)

# + ggplot2::coord_cartesian(xlim=c(-10, 10))
# ggplot2::coord_cartesian(xlim=c(-5, 5)

# with ggplot
ggplot(as.data.frame(resLFC_WT_BL_vs.pcry_BL),aes(log2FoldChange,-log(padj))) +
  geom_point() +
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rownames(top_WT_BL_vs.pcry_BL),]), colour = "blue") + 
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), colour = "red") +
  geom_text_repel(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), aes(label=anno[rocs,"id.symbol"]),colour = "red",hjust=-0.5, vjust=-1) +
  xlim(-5,5)
getwd()
ggsave("graphs/Vulcano ROCs-2.pdf",
       width = 10,
       height = 6)

## enhanced vulcano
library(airway)
  library(magrittr)

 library('DESeq2')

  dds <- DESeqDataSet(airway, design = ~ cell + dex)
  dds <- DESeq(dds, betaPrior=FALSE)
  res <- results(dds,
    contrast = c('dex','trt','untrt'))
  res <- lfcShrink(dds,
    contrast = c('dex','trt','untrt'), res=res, type = 'normal')
 EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    y = 'pvalue')

```

#### ---------------------------
#### COPY & PASTE from pCRY
#### Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}
outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
setwd(outdir)

dds <- readRDS(paste(outdir,"dds_23design.RDS",sep="/"))
anno <- readRDS(file=paste(outdir,"anno.RDS", sep="/"))

design(dds)
resultsNames(dds)
cbind(resultsNames(dds),c(rep(0,6)))

# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))

# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")

res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))

res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))

# Unterschiede zwischen versch. Belichtungen über beide Genotypen
# D vs. (BL+R)
res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))

# D vs. BL
res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))

# D vs. R
res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))

# WT vs. pCRY
res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))

## Make top lists:

topgenes <- function(res){
a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
a$id.symbol <- anno[rownames(a),"id.symbol"]
a$symbol <- anno[rownames(a),"geneSymbol"]
a <- a[order(a$baseMean, decreasing = T),]
  a$rank.bm <- seq(1:length(rownames(a)))
a <- a[order(a$padj, decreasing = F),]
  a$rank.padj <- seq(1:length(rownames(a)))
a <- a[order(abs(a$log2FoldChange), decreasing = T),]
  a$rank.l2FC <- seq(1:length(rownames(a)))
a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
  a <- a[order(a$rank.sum),]
print(a)
}

top.res1 <- topgenes(res=res1)
summary(res1)
top.res2 <- topgenes(res=res2)
summary(res2)
top.res3 <- topgenes(res=res3)
summary(res3)
top.res4 <- topgenes(res=res4)
summary(res4)
top.res5 <- topgenes(res=res5)
summary(res5)
top.res6 <- topgenes(res=res6)
summary(res6)
top.res7 <- topgenes(res=res7)
summary(res7)

top.all <- rbind(top.res1,
         top.res2,
         top.res3,
         top.res4,
         top.res5,
         top.res6,
         top.res7)
top.all <- top.all[order(top.all$rank.sum),]
head(top.all, n=185) %>% rownames() %>% unique() %>% length()
top.all <- top.all[unique(rownames(top.all)),]

# top.all <- head(top.all, n=100)

## Shrinkage
res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
res7_shrink <- lfcShrink(dds, res=res7, type="ashr")

```

### combine L2FC with padj
```{r, eval=FALSE}

res3_shrink
subset(mcols(dds), id.symbol =="SELU1") %>% rownames()
res3_shrink["Cre06.g263550",]

# DataFrame with 1 row and 5 columns
#                baseMean log2FoldChange     lfcSE      pvalue        padj
#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
# Cre06.g263550   3566.14        2.90397  0.167321 1.72033e-71 1.97322e-68

plot(res3_shrink$log2FoldChange,-log(res3_shrink$padj))
res3_shrink$logpadj <- -log(res3_shrink$padj, base=10)

-log(1.38379e-06,base = 10)

x=-2.5
60/(3^2)

y=-8*x^2 + 50
y

res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < res3_shrink$logpadj,TRUE,FALSE)
summary(res3_shrink$ishs)
subset(res3_shrink,ishs==TRUE & logpadj< 50 & abs(log2FoldChange)<2.5)
subset(res3_shrink,ishs==TRUE & logpadj< 50)

# for log2 -/+ 5 and padj 100 
res1_shrink$ishs <- ifelse((-20/3*(res1_shrink$log2FoldChange^2) + 60) < -log(res1_shrink$padj, base=10),TRUE,FALSE)
res2_shrink$ishs <- ifelse((-20/3*(res2_shrink$log2FoldChange^2) + 60) < -log(res2_shrink$padj, base=10),TRUE,FALSE)
res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < -log(res3_shrink$padj, base=10),TRUE,FALSE)
res4_shrink$ishs <- ifelse((-20/3*(res4_shrink$log2FoldChange^2) + 60) < -log(res4_shrink$padj, base=10),TRUE,FALSE)
res5_shrink$ishs <- ifelse((-20/3*(res5_shrink$log2FoldChange^2) + 60) < -log(res5_shrink$padj, base=10),TRUE,FALSE)
res6_shrink$ishs <- ifelse((-20/3*(res6_shrink$log2FoldChange^2) + 60) < -log(res6_shrink$padj, base=10),TRUE,FALSE)
res7_shrink$ishs <- ifelse((-20/3*(res7_shrink$log2FoldChange^2) + 60) < -log(res7_shrink$padj, base=10),TRUE,FALSE)

```

```{r, eval=FALSE}
outdir
figures <- paste(outdir,"figures",sep="/")
dir.create(figures)

```

### Colours
```{r, eval=FALSE}
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
```

### PCA (Panel B)
```{r, eval=FALSE}
library("PCAtools")
library(ggalt)

vsd <- vst(dds, blind=FALSE)
length(vsd)

dubs <- duplicated(mcols(vsd)$id.symbol)
length(dubs)

unique <- unique(mcols(vsd)$id.symbol)
length(unique)

vsd2 <- vsd[!duplicated(mcols(vsd)$id.symbol)]
rownames(vsd2) <- mcols(vsd2)$id.symbol

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
names(group.colors) <- levels(colData(dds)$condition)
group.colors

plotPCA(vsd2, intgroup="condition")
vst <- assay(vsd2)
p <- pca(vst, metadata = colData(dds), removeVar = 0.1)
screeplot(p, axisLabSize = 18, titleLabSize = 22)
biplot(p, colby = "condition")
biplot(p, showLoadings = TRUE,labSize = 5, pointSize = 5, sizeLoadingsNames = 5)

biplot(p,x= "PC1", y="PC3",
       lab = p$metadata$sample,
    colby = 'condition', colkey = group.colors,
    encircle = TRUE,
    encircleFill = TRUE,
    legendPosition = 'right',
    labSize = 4)
ggsave(paste(figures,"2023_12_PCA1-3.png",sep="/"),
  width = 6,
  height = 6)

biplot(p,x= "PC3", y="PC4",
       lab = p$metadata$sample,
    colby = 'condition', colkey = group.colors,
    encircle = TRUE,
    encircleFill = TRUE,
    legendPosition = 'right')
ggsave(paste(figures,"2023_12_PCA3-4.pdf",sep="/"),
  width = 12,
  height = 12)

pairsplot(p,colby = 'condition', colkey = group.colors)
ggsave(paste(figures,"2023_12_pairsplot.pdf",sep="/"),
  width = 12,
  height = 12)

plotloadings(p,
    rangeRetain = 0.01,
    labSize = 2.0,
    title = 'Loadings plot',
    subtitle = 'PC1, PC2, PC3, PC4, PC5',
    caption = 'Top 1% variables',
    drawConnectors = TRUE)
ggsave(paste(figures,"2023_12_loadings.pdf",sep="/"),
  width = 12,
  height = 12)

# Determine optimum number of PCs to retain
elbow <- findElbowPoint(p$variance)
elbow

anno["Cre01.g016600",]

 biplot(p,
    lab = p$metadata$condition,
    colby = 'condition', colkey = group.colors,
    hline = 0, vline = 0,
    encircle = TRUE, encircleFill = TRUE,
    legendLabSize = 16, legendIconSize = 8.0,
    legendPosition = 'right')

pairsplot(p,
    components = getComponents(p, c(1:6)),
    triangle = TRUE, trianglelabSize = 12,
    hline = 0, vline = 0,
    pointSize = 2,
    gridlines.major = FALSE, gridlines.minor = FALSE,
    colby = 'condition', colkey = group.colors,
    title = 'Pairs plot', plotaxes = FALSE,
    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))

```

### Counts
```{r, eval=FALSE}

# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Combined counts table TFs

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
max_val <- 1.0*max(all_counts$count)

# Plot
gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  scale_y_continuous(trans = "log2")
gcounts
ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
  width = 12,
  height = 8)

```

### Volcanos
### dark
```{r, eval=FALSE}

# dark
res <- res1_shrink
n <- 'WT vs. pCRY in dark'
nf <- 'figures/Volcano_dark.png'
l <- nrow(top.res1)
pcol <- "black"
lcol <- colours[1+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evd <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-60),
    FCcutoff = 3,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evd

```

### blue
```{r, eval=FALSE}

# blue
res <- res2_shrink
n <- 'WT vs. pCRY in blue'
nf <- 'figures/Volcano_blue.png'
l <- nrow(top.res2)
pcol <- colours[3]
lcol <- colours[3+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 17,
                ifelse(res$padj < 1e-100,17,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evb <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evb

```

### red
```{r, eval=FALSE}

# red
res <- res3_shrink
n <- 'WT vs. pCRY in red'
nf <- 'figures/Volcano_r.png'
l <- nrow(top.res3)
pcol <- colours[5]
lcol <- colours[5+1]

rownames(res) <- mcols(dds)$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]

shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
                ifelse(res$padj < 1e-100,6,16))
summary(is.na(shape))
# shape[is.na(shape)] <- 2
names(shape)[shape == 6] <- 'outliner'
names(shape)[shape == 16] <- 'in range'

res$log2FoldChange[res$log2FoldChange > 5] <- 5
res$log2FoldChange[res$log2FoldChange < -5] <- -5
res$padj[res$padj < 1e-100] <- 1e-100
summary(res$padj < 1e-100)

keyvals <- ifelse(
    res$ishs == TRUE, pcol,
      'grey')

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == pcol] <- 'highly significant'
  names(keyvals)[keyvals == 'grey'] <- 'other'

evr <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = rownames(res),
    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
    # selectLab = anno_tf$geneSymbol,
    # boxedLabels = TRUE,
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = TRUE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 10,
    colCustom = keyvals,
    xlim = c(-5.1, 5.1),
    ylim = c(0, 101),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("DE genes:",l),
    # sub = "SVF",
    pCutoff = 10^(-50),
    FCcutoff = 2.5,
   shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
                   'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
evr

#ggsave(file="figures/Vulcano_rtest.pdf",
#       width = 12,
#       height = 12)

```

### all
```{r, eval=FALSE}

ga <- ggarrange(evb,
                evd+rremove("ylab"),
                evr+rremove("ylab"),
                ncol = 3, nrow = 1,common.legend = TRUE, legend = "bottom",
                widths = c(1.05,1,1))

ggsave(ga, file="figures/Vulcano_all.pdf",
       width = 12,
       height = 8)

# + ggplot2::coord_cartesian(xlim=c(-10, 10))
# ggplot2::coord_cartesian(xlim=c(-5, 5)

# with ggplot
ggplot(as.data.frame(resLFC_WT_BL_vs.pcry_BL),aes(log2FoldChange,-log(padj))) +
  geom_point() +
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rownames(top_WT_BL_vs.pcry_BL),]), colour = "blue") + 
  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), colour = "red") +
  geom_text_repel(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), aes(label=anno[rocs,"id.symbol"]),colour = "red",hjust=-0.5, vjust=-1) +
  xlim(-5,5)
getwd()
ggsave("graphs/Vulcano ROCs-2.pdf",
       width = 10,
       height = 6)

## enhanced vulcano
library(airway)
  library(magrittr)

 library('DESeq2')

  dds <- DESeqDataSet(airway, design = ~ cell + dex)
  dds <- DESeq(dds, betaPrior=FALSE)
  res <- results(dds,
    contrast = c('dex','trt','untrt'))
  res <- lfcShrink(dds,
    contrast = c('dex','trt','untrt'), res=res, type = 'normal')
 EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    y = 'pvalue')

```

### Heatmap

### plot heatmap
```{r, eval=FALSE}
library("pheatmap")
ntd <- normTransform(dds)
top.all.hm <- head(top.all, n=100)
df <- assay(ntd)[rownames(top.all.hm),]
rownames(df) <- mcols(dds)[rownames(df),"id.symbol"]

# anno_col <- as.data.frame(colData(dds)[,c("condition","treatment","genotype")])
anno_col <- data.frame(condition = dds$condition, treatment = dds$treatment)
rownames(anno_col) <- rownames(colData(dds))

# column order
anno_col <- arrange(anno_col, match(anno_col$condition, levels(anno_col$condition)))
levels(anno_col$condition)
df <- df[,rownames(anno_col)]

annot_colors=list(condition=c(Cancer="#F0978D",Healthy="#63D7DE"))
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)

mycolors2 <- list(genotype = c("#525252","#969696"),
     treatment = c("#525252","#1F78B4","#E31A1C"),
     condition = colours)

names(mycolors2$genotype) <- levels(anno_col$genotype)
names(mycolors2$treatment) <- levels(anno_col$treatment)
names(mycolors2$condition) <- levels(anno_col$condition)
mycolors2

xx <- pheatmap(df, cluster_rows=TRUE, show_rownames=TRUE,
         cluster_cols=FALSE,
         annotation_col=anno_col,
         color=colorRampPalette(c("navy", "white", "red"))(50),
         annotation_colors=mycolors2
         )

ggsave("figures/2023_12_Heatmap_top.png",plot=xx,
       width = 5,
       height = 15)

```

### Complex Heatmap
```{r, eval=FALSE}

df
genes <- anno_tf

# collapse replicates
colData(dds)
dds_heat <- collapseReplicates(dds, dds$condition,dds$sample)
dds_heat$sample <- factor(dds_heat$sample, levels=c("WT-D1","pcry-D1","WT-BL1","pcry-BL1","WT-R1","pcry-R1"))
colData(dds_heat)
dds_heat$runsCollapsed
colnames(dds_heat)
matchFirstLevel <- dds_heat$sample == levels(dds_heat$sample)[1]
stopifnot(all(rowSums(counts(dds_heat[,matchFirstLevel])) == counts(dds_heat[,1])))

mat <- counts(dds_heat)
length(rownames(counts(dds_heat)))
length(rownames(df))
rownames(mat) <- mcols(dds_heat)$id.symbol
length(rownames(mat))

mat <- mat[df,]

position <- mat
l <- length(rownames(mat))
pos <- data.frame(genes = rownames(mat),pos = 1:l)
rownames(pos) <- rownames(mat)
pos <- pos[genes,2]

# mat <- cbind(res,counts(dds_heat))
# mat <- subset(mat,!SYMBOL2=="" & SYMBOL2 %in% df$SYMBOL2)
# rownames(mat) <- mat$SYMBOL2
# mat <- mat[,c(12:15)]
heat <- t(scale(t(mat)))
Heatmap(heat)

summary(heat)
myCol <- colorRampPalette(c('brown3', 'yellow2', 'darkolivegreen3'))(100)
myBreaks <- seq(-1.5, 1.5, length.out = 100)

ann <- data.frame(condition = colData(dds_heat)[,"condition"],
  stringsAsFactors = FALSE)

colAnn <- HeatmapAnnotation(
  df = ann,
  which = 'col', # 'col' (samples) or 'row' (gene) annotation?
  na_col = 'white', # default colour for any NA values in the annotation data-frame, 'ann'
  # col = colours,
  annotation_height = 0.6,
  annotation_width = unit(1, 'cm'),
  gap = unit(1, 'mm'))

boxplotCol <- HeatmapAnnotation(
  boxplot = anno_boxplot(
    heat,
    border = FALSE,
    gp = gpar(fill = '#CCCCCC'),
    pch = '.',
    size = unit(2, 'mm'),
    axis = TRUE,
    axis_param = list(
      gp = gpar(fontsize = 12),
      side = 'left')),
  annotation_width = unit(c(2.0), 'cm'),
  which = 'col')

boxplotRow <- HeatmapAnnotation(
  boxplot = row_anno_boxplot(
    heat,
    border = FALSE,
    gp = gpar(fill = '#CCCCCC'),
    pch = '.',
    size = unit(2, 'mm'),
    axis = TRUE,
    axis_param = list(
      gp = gpar(fontsize = 12),
      side = 'top')),
  annotation_width = unit(c(2.0), 'cm'),
  which = 'row')

genelabels <- rowAnnotation(
  Genes = anno_mark(
    at = pos,
    labels = genes,
    labels_gp = gpar(fontsize = 15, fontface = 'bold'),
    padding = 0.75),
  width = unit(2.0, 'cm') +

    max_text_width(
      rownames(heat)[seq(1, nrow(heat), 40)],
      gp = gpar(fontsize = 10,  fontface = 'bold')))

# Clusters
# pamClusters <- cluster::pam(heat, k = 4) # pre-select k = 4 centers
pamClusters <- cluster::pam(heat, k= 8) # pre-select k = 4 centers

pamClusters$clustering <- paste0('Cluster ', pamClusters$clustering)
pamClusters$clustering <- factor(pamClusters$clustering,
                                 levels = c('Cluster 1', 'Cluster 2', 'Cluster 3', 'Cluster 4','Cluster 5', 'Cluster 6'))

hmap <- Heatmap(heat,

                # split the genes / rows according to the PAM clusters
                split = pamClusters$clustering,
                cluster_row_slices = TRUE,

                name = 'Gene\nZ-\nscore',

                col = colorRamp2(myBreaks, myCol),

                # parameters for the colour-bar that represents gradient of expression
                heatmap_legend_param = list(
                  color_bar = 'continuous',
                  legend_direction = 'vertical',
                  legend_width = unit(8, 'cm'),
                  legend_height = unit(5.0, 'cm'),
                  title_position = 'topcenter',
                  title_gp=gpar(fontsize = 12, fontface = 'bold'),
                  labels_gp=gpar(fontsize = 12, fontface = 'bold')),

                # row (gene) parameters
                cluster_rows = TRUE,
                show_row_dend = TRUE,
                row_title = 'Statistically significant genes',
                row_title_side = 'left',
                row_title_gp = gpar(fontsize = 12,  fontface = 'bold'),
                row_title_rot = 90,
                show_row_names = FALSE,
                row_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                row_names_side = 'left',
                row_dend_width = unit(25,'mm'),

                # column (sample) parameters
                cluster_columns = TRUE,
                show_column_dend = TRUE,
                column_title = '',
                column_title_side = 'bottom',
                column_title_gp = gpar(fontsize = 12, fontface = 'bold'),
                column_title_rot = 0,
                show_column_names = TRUE,
                column_names_gp = gpar(fontsize = 10, fontface = 'bold'),
                column_names_max_height = unit(10, 'cm'),
                column_dend_height = unit(25,'mm'),

                # cluster methods for rows and columns
                clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
                clustering_method_columns = 'ward.D2',
                clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
                clustering_method_rows = 'ward.D2',

                # specify top and bottom annotations
                top_annotation = colAnn)
                # bottom_annotation = boxplotCol)

pdf("Heatmap2.pdf", height = 18, width = 6)
draw(hmap + genelabels,
     heatmap_legend_side = 'left',
     annotation_legend_side = 'right',
     row_sub_title_side = 'left')
dev.off()

```

# Export

```{r, eval=FALSE}

library(writexl)

# combine results

res <- res_WT_BL_vs.pcry_BL
res1 <- res_WT_D_vs.pcry_D
res2 <- res_WT_BL_vs.pcry_BL
res3 <- res_WT_R_vs.pcry_R
res4 <- res_WT_D_vs.WT_BL
res5 <- res_WT_D_vs.WT_R

colnames(anno)

res_exp <- data.frame(
  # general
  "gene_id" = rownames(dds),
  "gene_name" = anno[rownames(dds),"geneSymbol"],
  "Alias" = anno[rownames(dds),"prev.symbols"],
  "id.symbol" = anno[rownames(dds),"id.symbol"],
 "Description" = anno[rownames(dds),"Description"],
 "Comments" = anno[rownames(dds),"Comments"],
 "TargetP" = anno[rownames(dds),"TargetP"],
 "Predalgo" = anno[rownames(dds),"Predalgo"],
 "baseMean" = res$baseMean,
 # results 1
 "L2FC.WT_D_vs.pcry_D" = res1$log2FoldChange,
 "pvalue.WT_D_vs.pcry_D" = res1$pvalue,
 "padj.WT_D_vs.pcry_D" = res1$padj,
  # results 2
 "L2FC.WT_BL_vs.pcry_BL" = res2$log2FoldChange,
 "pvalue.WT_BL_vs.pcry_BL" = res2$pvalue,
 "padj.WT_BL_vs.pcry_BL" = res2$padj,
  # results 3
 "L2FC.WT_R_vs.pcry_R" = res3$log2FoldChange,
 "pvalue.WT_R_vs.pcry_R" = res3$pvalue,
 "padj.WT_R_vs.pcry_R" = res3$padj,
  # results 4
 "L2FC.WT_D_vs.WT_BL" = res4$log2FoldChange,
 "pvalue.WT_D_vs.WT_BL" = res4$pvalue,
 "padj.WT_D_vs.WT_BL" = res4$padj,
  # results 5
 "L2FC.WT_D_vs.WT_R" = res5$log2FoldChange,
 "pvalue.WT_D_vs.WT_R" = res5$pvalue,
 "padj.WT_D_vs.WT_R" = res5$padj,
  counts(dds, normalized=TRUE))
res_exp

outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"

write_xlsx(data.frame(res_exp),
           paste(outdir,"2023_08 P3044 results.xlsx",sep="/"))

# Counts only
counts <- round(counts(dds, normalized=TRUE))
colnames(counts)
colData(dds)$sample.name <- paste(colData(dds)$condition,colData(dds)$repetition,sep="_")
colnames(counts) == colData(dds)$sample.id
colnames(counts) <- colData(dds)$sample.name

write.csv(counts, file.path(outdir,"Counts.csv"))

Exp_design <- data.frame(names = colData(dds)$sample.name,
                         genotype = colData(dds)$genotype,
                         treatment = colData(dds)$treatment,
                         condition = colData(dds)$condition) %>% t()

Exp_design <- matrix(colData(dds)$sample.name,
                     colData(dds)$genotype,
                     colData(dds)$treatment,
                    colData(dds)$condition)

rownames(Exp_design)[1] <- {1}

write.table(Exp_design,file.path(outdir,"Exp_design.csv"), sep=",",  col.names=FALSE)

```

```{r}
sessionInfo()
```