### 2024-10-19 pub_figures/manuscript-figures.Rmd
Venn + Grouping

---
title: "figures_manuscript"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
        })
date: "2024-10-17"
editor_options: 
  chunk_output_type: inline
---

```{r knit, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

```

BiocManager::install()

# run prepare_data scripts
```{r run_start_up, include=FALSE, results='hide'}
library(knitr)
source(purl("/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/git_Chlamy_RNAseq_aCRY/2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd", output = tempfile()), local = knitr::knit_global())

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")

```

# Fig 1: PCA
```{r 1_pca, fig.show='hold', out.width="50%", dpi=300, results='hide'}

fig <- "Fig1"

# group
for (dds in dds_list) {
vsd <- vst(dds, blind=FALSE)
# colData(dds)
### PCA with top 500 genes with highest row variance 
pcaData <- plotPCA(vsd, intgroup=c("treatment", "genotype","condition"), returnData=TRUE,)
percentVar <- round(100 * attr(pcaData, "percentVar"))

g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(size=4, stroke = 1) +
  scale_shape_manual(values = c(16,21)) + 
  labs(title = "PCA plot") +
  ylab(paste0("PC1: ",percentVar[1],"% variance")) +
  xlab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment) +
  # coord_fixed() +
  scale_x_reverse() +
  theme_bw() +
  removeGrid(x=T, y=T)
}

# pCRY
dds <- dds_list[["pcry"]]
vsd <- vst(dds, blind=FALSE)

pcaData <- plotPCA(vsd, intgroup=c("treatment", "genotype","condition"), returnData=TRUE,)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g1 <- ggplot(pcaData, aes(PC2, PC1, color=treatment, shape=genotype)) +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(size=4, stroke = 1) +
  scale_shape_manual(values = c(16,21)) + 
  labs(title = "PCA plot: pCRY") +
  ylab(paste0("PC1: ",percentVar[1],"% variance")) +
  xlab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment) +
  # coord_fixed() +
  scale_x_reverse() +
  scale_y_reverse() +
  theme_bw() +
  removeGrid(x=T, y=T)
g1

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_PCA.pdf",sep=""), plot = g1,
width = 6,
height = 6)

# aCRY
dds <- dds_list[["acry"]]
vsd <- vst(dds, blind=FALSE)
pcaData <- plotPCA(vsd, intgroup=c("treatment", "genotype","condition"), returnData=TRUE,)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g2 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(size=4, stroke = 1) +
  scale_shape_manual(values = c(16,21)) + 
  labs(title = "PCA plot: aCRY") +
  ylab(paste0("PC1: ",percentVar[1],"% variance")) +
  xlab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment) +
  # coord_fixed() +
  scale_x_reverse() +
  theme_bw() +
  removeGrid(x=T, y=T)
g2

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_PCA.pdf",sep=""), plot = g2,
width = 6,
height = 6)

```

# Fig 2: Counts
```{r Counts, echo=FALSE, fig.width=10}

fig <- "Fig2"

dds <- dds_list[[1]]
for (dds in dds_list){
# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts_tf1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  labs(title = "ROC genes") + 
  theme_bw() +
  removeGrid(x=T, y=T) +
    geom_vline(xintercept=seq(1,length(levels(all_counts$Gene))-1,1)+.5,color="grey") +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_tf1 %>% print()

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_ROCs.pdf",sep=""), plot = gcounts_tf1,
width = 12,
height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Combined counts table TFs

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# Plot
max_val <- 1.0*max(all_counts$count)
gcounts_phot1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  labs(title = "Photoreceptor genes") + 
  theme_bw() +
  removeGrid(x=T, y=T) +
  geom_vline(xintercept=seq(1,length(levels(all_counts$Gene))-1,1)+.5,color="grey") +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_phot1 %>% print()

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_Phots.pdf",sep=""), plot = gcounts_phot1,
width = 12,
height = 8)

# print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))

# single gene
    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
    colnames(d)[1] <- "counts"

gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  # geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    theme_bw() +
  removeGrid(x=T, y=T) +
    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +

    labs(title = paste("ROC40"," in ",colData(dds)$experiment[1],sep = "")) +
 theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
        axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))

gcounts %>% print()

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_ROC40.pdf",sep=""), plot = gcounts,
width = 6,
height = 5)
}

# ggsave(paste(pub_figures,"ROC40.pdf",sep="/"), plot = gcounts,
#   width = 5,
#   height = 4)

```

# Fig 3: Volcanos
```{r volcanos, fig.height=6, fig.width=10}

fig <- "Fig3"

dds <- dds_list[["pcry"]]

res_ashr_list %>% names()
res_ashr_list[[1]] %>% names()

res <- res_ashr_list$pcry$pcry_R.vs.WT_R
res_n <- res_list$pcry$pcry_R.vs.WT_R

# of shrinked results
total <- subset(res, padj< 0.05) %>% nrow()
up <- subset(res, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# of "true" results
total <- subset(res_n, padj< 0.05) %>% nrow()
up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# points outside the grid
subset(res, padj < 10^-50 | log2FoldChange > 6 | log2FoldChange < -6)
#                 baseMean log2FoldChange     lfcSE      pvalue        padj
#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
# Cre16.g681750 2907.4406       -3.51641   0.22216 2.99151e-57 4.79270e-53
# Cre17.g802135   46.8587       36.40049   7.35075 1.00156e-08 7.64096e-06

res["Cre16.g681750","padj"] <- 10^-50
res["Cre17.g802135","log2FoldChange"] <- 6

mcols(dds_list[["pcry"]]) %>% nrow()
res %>% nrow()

volcano_red <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[5]),
    title = "red-light, pCRY vs. WT",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 40,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)
# volcano_red

# blue

res_ashr_list %>% names()
res_ashr_list[[1]] %>% names()

res <- res_ashr_list$pcry$pcry_BL.vs.WT_BL
res_n <- res_list$pcry$pcry_BL.vs.WT_BL

# of "true" results
total <- subset(res_n, padj< 0.05) %>% nrow()
up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# points outside the grid

pmax <- 10^-50
l2FCmax <- 6
subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
# res[res$padj < pmax,]$padj <- pmax

subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

mcols(dds_list[["pcry"]]) %>% nrow()
res %>% nrow()

volcano_blue <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[3]),
    title = "blue-light, pCRY vs. WT",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 40,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)
# volcano_blue

# dark

res_ashr_list %>% names()
res_ashr_list[[2]] %>% names()

res <- res_ashr_list$pcry$pcry_D.vs.WT_D
res_n <- res_list$pcry$pcry_D.vs.WT_D

# of "true" results
total <- subset(res_n, padj< 0.05) %>% nrow()
up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# points outside the grid

pmax <- 10^-50
l2FCmax <- 6
subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
res[subset(res, padj < pmax) %>% rownames(),]$padj <- pmax

subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

mcols(dds_list[["pcry"]]) %>% nrow()
res %>% nrow()

volcano_dark <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[1]),
    title = "dark, pCRY vs. WT",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 40,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)
# volcano_dark

volcanos_all <- volcano_dark + volcano_blue + volcano_red +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'none', axis.title=element_text(size=12))
volcanos_all

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
width = 10,
height = 6)

```

# Fig 4: Overlaps
## Fig 4a: Venns
```{r overlap, eval=TRUE}
fig <- "Fig4"

res_list[["pcry"]] %>% names()
#  "pcry_D.vs.WT_D"       "pcry_BL.vs.WT_BL"     "pcry_R.vs.WT_R"      
DEGs <- list(dark = res_list$pcry$pcry_D.vs.WT_D %>% subset(padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )),
             blue = res_list$pcry$pcry_BL.vs.WT_BL %>% subset(padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )),
             red =res_list$pcry$pcry_R.vs.WT_R %>% subset(padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )))
DEGs <- lapply(DEGs,data.frame)
DEGs_genes <- lapply(DEGs,rownames)

venn.ol <- calculate.overlap(DEGs_genes)
venn.ol %>% lapply(length)

input_list <- DEGs_genes

plt <- venn.diagram(
    x = input_list,
    inverted=TRUE,
    # total.population = TRUE,
    filename = NULL,
    fontfamily ="Arial",
    lwd = 2,
    lty = 'blank',
    fill = group.colors[c(2,4,6)],
    category.names = paste0(names(input_list),"\n(",lapply(input_list,length),")"),
#    cat.col=cols2d[c(1,2)],
    cat.fontface = "bold",
    cat.fontfamily = "arial",
#    cat.pos = c(+30,-30),
#    cat.dist = c(0.12, 0.12),
    disable.logging = TRUE
)

wrap_elements(plt) # + plot_annotation(caption = paste0("pCRY"))

```

## Fig 4b: Groups
```{r groups, eval=FALSE}

res_core <- list(dark = res_list$pcry$pcry_D.vs.WT_D,
             blue = res_list$pcry$pcry_BL.vs.WT_BL,
             red =res_list$pcry$pcry_R.vs.WT_R) %>% lapply(data.frame)

DEGs_comb <- bind_cols(res_core[[1]],res_core[[2]][,c(2,6)],res_core[[3]][,c(2,6)])
colnames(DEGs_comb)[c(2,6,8,9,10,11)] <- c("l2FC.dark","padj.dark","l2FC.blue","padj.blue","l2FC.red","padj.red")

DEGs_genes["dark"]
blue_red_genes <- unlist(DEGs_genes[c("blue","red")]) %>% unique()

non-sig <- data.frame(x = c(-9, 0, 9), y = c(20, 0.5, 20))

DEGs_comb$sig

gg_all <- ggplot(DEGs_comb, aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
  geom_point(shape=21) +
  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
  geom_text_repel() +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_all

gg_br <- ggplot(DEGs_comb[blue_red_genes,], aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
  geom_point(shape=21) +
  geom_text_repel() +
  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
  theme_bw() +
  removeGrid(x=T, y=T)
gg_br

red_specific <- DEGs_comb %>% subset(l2FC.blue < 1 & l2FC.blue > -1 & (l2FC.red > 0.5 | l2FC.red < -0.5) & padj.red < 0.05) %>% arrange(l2FC.red)

goi <- "Cre03.g192050"

# single gene
d <- plotCounts(dds, gene=goi, intgroup=c("condition","experiment","genotype","treatment"), main=s,returnData=TRUE)
colnames(d)[1] <- "counts"

gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  # geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    theme_bw() +
    removeGrid(x=T, y=T) +
    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +
    labs(title = paste(mcols(dds)[goi,]$symbol," in ",colData(dds)$experiment[1],sep = "")) +
    theme(axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
    axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
    axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))
gcounts

```

# Fig. X
```{r 3_groups, eval=FALSE, fig.show='hold', fig.width=6, fig.height=4, dpi=300, out.width="50%"}

p_new2 <- ggplot(deg_table, aes(x=XX.log2FC, y=XY.log2FC,color=group, fill=group)) + 
  geom_polygon(data = xyup, aes(x=x,y=y),color=cols2l[2],fill=cols2l[2], alpha=0.2) +
  geom_polygon(data = xydo, aes(x=x,y=y),color=cols2l[2],fill=cols2l[2], alpha=0.8) +
  geom_polygon(data = xxdo, aes(x=x,y=y),color=cols2l[1],fill=cols2l[1], alpha=0.8) +
  geom_polygon(data = xxup, aes(x=x,y=y),color=cols2l[1],fill=cols2l[1], alpha=0.2) +
  annotate("text", x= text$x,y= text$y,label = paste(text$label,text$size,sep="\n"), color=text$color,size=3 , fontface="bold") +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(shape=21) + 
  scale_color_manual(values=alpha(c(cols2d[1],cols2d[1], cols2d[2],cols2d[2],rep(colsp,2),rep(colsap,2)),0.8)) +
  scale_fill_manual(values=alpha(c(cols2d[1],cols2d[1], cols2d[2],cols2d[2],rep(colsp,2),rep(colsap,2)),0.3)) +
#  scale_color_manual(values = text$color) +
  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
  theme_bw() +
  removeGrid(x=T, y=T)

p_new2 + scale_fill_manual(values=c(cols2d[1],cols2d[1], cols2d[2],cols2d[2],alpha(c(rep(colsp,2),rep(colsap,2)),0.3)))

```